### 2017-12-08 doc/presentation/presentation.tex
Skeleton for presentation

\documentclass[9pt]{beamer}
\usecolortheme{beaver}
\setbeamertemplate{navigation symbols}{}
\setbeamercolor{block title}{bg=white!95!black,fg=black}
\defbeamertemplate{description item}{align left}{\insertdescriptionitem\hfill}
    \setbeamertemplate{description item}[align left]
\usepackage{tikz}

\usepackage[outputdir=build]{minted}
\usepackage{tcolorbox}
\tcbuselibrary{minted,skins}

% Patch to fix https://github.com/T-F-S/tcolorbox/issues/12
\makeatletter
\def\tcb@minted@input@listing#1#2#3#4{%
  \edef\temp@a{#4}%
  \ifx\temp@a\@empty%
  \else%
    \toks@=\expandafter{#4}%
    \edef\tcb@temp{\noexpand\usemintedstyle{\the\toks@}}%
    \tcb@temp%
  \fi%
  \toks@=\expandafter{#1}%
  \edef\tcb@temp{\noexpand\inputminted[\the\toks@]}%
  \IfFileExists{\minted@outputdir#3}%
    {\tcb@temp{#2}{\minted@outputdir#3}}%
    {\tcb@temp{#2}{#3}}%
}
\makeatother

\newtcblisting{ccode}[2][]{%
  listing engine=minted,
  minted language=#2,
  minted options={breaklines,breakanywhere,fontsize=\scriptsize,gobble=8},
  listing only,
  before skip=0pt,
  after skip=8pt,
  left skip=0pt,
  right skip=0pt,
  size=fbox,
  sharp corners,
  %colframe=white!75!black,
  colframe=white,
  boxrule=0pt,
  frame hidden,
  #1
}

\newtcbinputlisting[]{\inputcode}[3][]{%
  listing engine=minted,
  minted language=#2,
  minted options={breaklines,breakanywhere,fontsize=\scriptsize,#1},
  listing file={#3},
  listing only,
  size=fbox,
  before skip=0pt,
  after skip=10pt,
  left skip=0pt,
  right skip=0pt,
  sharp corners,
  %colframe=white!75!black,
  colframe=white,
  boxrule=0pt,
  frame hidden
}

\title{Small Sparse Matrix Multiplications\\ on Knights Landing}
\author{Nathan Brei}
\institute{Technical University of Munich}
\date\today

\begin{document}
\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Motivation}

  \begin{block}{SeisSol}
    Simulates earthquakes, specifically, dynamic rupture processes and seismic wave propagation over an unstructured tetrahedral mesh. 
  \end{block}

  \begin{block}{ADER-DG}
    A numerical method which uses Discontinuous Galerkin discretization in space and ADER discretization in time. It avoids a global system matrix, using local element matrix operations instead.
  \end{block}

  \begin{block}{Small Sparse Matrix Multiplications}
    The compute kernels resemble the operation shown below, applied recursively. Different sparsity patterns are used, but they are all known at compile time.
  \end{block}

  \centering
  \includegraphics[height=3cm]{images/seissol_visc.png}

\end{frame}

\begin{frame}
  \frametitle{Knights Landing}
  \begin{block}{KNL is the 2nd-generation Intel Xeon Phi many-core processor}
    \begin{itemize}
    \item A standalone processor (not just a coprocessor or GPU) which supports the entire x86 instruction set along with AVX512 extensions.
    \item Organized into a 2D grid of up to 36 tiles, each containing 2 cores and a 1MB shared L2 cache
    \item 4 threads/core $\implies$ up to 288 threads/processor
    \item 32 vector registers holding 8 doubles/register
    \item Fused-multiply-add instruction $(=:FMA)$ with mask registers and implicit broadcast
    \item 2 vpus/core, 16 double ops/vpu/cycle, 1.3e9 cycles/sec \\ $\implies$ 3 double TFlop peak performance

    \end{itemize}
  \end{block}
  \begin{block}{KNL greatly prefers large dense matrix operations to small sparse ones}
    \begin{itemize}
    \item Scalar operations limited to 1/8 peak performance
    \item Bottleneck with vector stores: 64B/cycle
    \item Bottleneck with instruction cache, decoder: min(16B/cycle, 2 instructions)
    \item \textbf{Out-of-order execution pipeline can only issue 2 instructions per cycle}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Past approaches used in SeisSol}
  \begin{block}{Standard compressed-sparse-column multiplication}
    \begin{itemize}
      \item[$+$] Generality.
      \item[$+$] Algorithm uses a (small) constant number of instructions.
      \item[$-$] Index lookups require indirect memory access, causing lots of problems for the execution pipeline.
      \item[$-$] $sparse \times sparse$ is fundamentally scalar. (However, BSR can be vectorized.)
      \item[$-$] Disregards the fact that the sparsity pattern is known at compile time.
    \end{itemize}
  \end{block}

  \begin{block}{Sparsity pattern unrolled into instruction stream}
    \begin{itemize}
    \item[$+$] Index lookups removed completely
    \item[$+$] $dense \times sparse$ case can be vectorized perfectly 
    \item[$-$] Generates one FMA instruction for every nonzero
    \item[$-$] (Implementation) Memory access pattern is suboptimal on KNL
    \item[$-$] (Implementation) Intel compiler struggles to generate efficient assembly
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Present and future approaches}
  \begin{block}{Dense matrix multiplication}
    \begin{itemize}
    \item[$+$] Memory access patterns and register blockings optimized for KNL
    \item[$+$] (Implementation) Code generator emits assembly directly 
    \item[$-$] Filling in zero entries wastes both flops and memory bandwidth
    \end{itemize}
  \end{block}

  \begin{block}{My approach}
    \begin{itemize}
    \item The three existing algorithms each encounter different constraints and bottlenecks.
    \item Think of them as poles in a design space. Explore algorithms which combine their features in order to finesse around the bottlenecks.
    \item Speed up applications such as SeisSol, or demonstrate why it can't be done.
    \item Create a code generator which autotunes a sparse gemm algorithm for a reasonably arbitrary problem.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{The dense outer-product microkernel}
  \begin{minted}[fontsize=\footnotesize]{gas}
# Load C register block
vmovapd 0(%rdx),  %zmm16                                 # load C[:,0]
vmovapd 64(%rdx), %zmm17                                 # load C[:,1]
vmovapd 128(%rdx), %zmm18                                # load C[:,2]
# ...

# Load A register block
vmovapd 0(%rdi), %zmm0                                   # load A[:,0]
vmovapd 64(%rdi), %zmm1                                  # load A[:,1]
# ...

# Outer product of A[:,0] * B[0,:]
vfmadd231pd 0(%rsi) {1to8}, %zmm0, %zmm16                # C[:,0] += A[:,0] .* B[0,0]
vfmadd231pd 0(%rsi,%r15,1) {1to8}, %zmm0, %zmm17         # C[:,1] += A[:,0] .* B[0,1]
vfmadd231pd 0(%rsi,%r15,2) {1to8}, %zmm0, %zmm18         # C[:,2] += A[:,0] .* B[0,2]
# ...

# Outer product of A[:,1] * B[1,:]
vfmadd231pd 8(%rsi) {1to8}, %zmm1, %zmm16                # C[:,0] += A[:,1] .* B[1,0]
vfmadd231pd 8(%rsi,%r15,1) {1to8}, %zmm1, %zmm17         # C[:,0] += A[:,1] .* B[1,1]
vfmadd231pd 8(%rsi,%r15,2) {1to8}, %zmm1, %zmm18         # C[:,0] += A[:,1] .* B[1,2]
# ...
# More outer products ...

# Store C register block
vmovapd %zmm16, 0(%rdx)                                  # store C[:,0]
vmovapd %zmm17, 64(%rdx)                                 # store C[:,1]
vmovapd %zmm18, 128(%rdx)                                # store C[:,2]
# ...

  \end{minted}
\end{frame}

%\begin{frame}
%  \frametitle{Cartoon of the dense outer-product microkernel}
%\end{frame}

\begin{frame}
  \frametitle{The sparse outer-product microkernel}

  \begin{block}{Idea: Build off of the state of the art for small dense gemm}
    \begin{itemize}
    \item Assume A is dense, B is sparse
    \item Pack B into a virtual compressed-sparse-row or column format
    \item Remove all FMA and load instructions which are not needed
    \item Make this the basic building block for more complicated algorithms 
    \end{itemize}
  \end{block}
  \begin{block}{Limitations}
    \begin{itemize}
    \item Sparsity pattern must be known at compile time
    \item Pattern must fit within vector registers: $m \leq 8;  n \leq 30$
    \item (Dense $\times$ Sparse) $\neq$ (Sparse $\times$ Dense) due to vectorization
    \item Choice of row-major vs column-major format determines which of the two vectorizes nicely: $(AB)^T = B^T A^T$
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Fully unrolled dense x sparse}
  \begin{columns}[t]%[onlytextwidth]
    \begin{column}{0.5\textwidth}
      Generate sparse microkernels for each (k,n) block of B, and unroll these into the instruction stream.
      \begin{itemize}
      \item[$+$] Supports arbitrary matrix sizes
      \item[$+$] Dense-style register blocking for a dense $\times$ sparse algorithm.
      \item[$+$] Does not fill in any zeros -- no wasted flops

      \item[$-$] Generates an FMA instruction for each nonzero
      \item[$-$] These FMA instructions are 10B, too large to stream
      \item[$-$] Irregular location of blocks in memory prevents using tricks to reduce instruction size
      \item[$-$] Total number of nonzeros is limited by size of instruction cache.

      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{ccode}[]
        {text}
        loop over m blocks:

           unroll loop over n blocks:

              load C block into registers

              unroll loop over k blocks:

                 load A block into registers
                 microkernel for current B block
                 move A right 1 block
                 move B down 1 block

              store C block from registers
              move A to far left
              move B to top + right 1 block
              move C right 1 block

           move A down 1 block
           move C down 1 block
      \end{ccode}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Performance of fully unrolled dense x sparse}
  Matrix sizes: $m=128, n=28, k=128$.
  Microkernel block sizes: $bm=8, bn=28, bk=4$.
  B is gradually filled with randomly placed nonzeros.

  \begin{columns}[t]
    \begin{column}{0.6\textwidth}
      \centering
      \includegraphics[width=6.5cm]{images/fig2.pdf}
    \end{column}
    \begin{column}{0.6\textwidth}
      \centering
      \includegraphics[width=6.5cm]{images/fig3.pdf}
    \end{column}
  \end{columns}

  \begin{itemize}
    \item For $nnz < 3100$, time-to-solution scales linearly.
    \item For $nnz > 3200$, the FMAs completely fill the instruction cache (regardless of matrix dimensions!)
    \item Fully unrolled sparse algorithm outperforms dense ``nonzero'' GFlops, but always underperforms dense ``hardware'' GFlops

  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Improvement: Tiled Sparse}
        Assume there exists a sparse pattern which tiles perfectly over B.
        \begin{itemize}
        \item[$+$] We can reuse a single microkernel, reducing the number of instructions
        \item[$+$] We can visit blocks in a loop efficiently, using pointer tricks to reduce instruction size.
        \item[$-$] This regularity assumption is likely too strong.
        \item[$-$] In practice, the sparsity pattern would often become dense.
        \end{itemize}

      \[
      \left[
          \begin{array}{c c c | c c c }
          0 &   &   & 10 &   &   \\
          1 & 2 &   & 11& 12&   \\
            & 3 & 4 &   & 13& 14  \\
          \hline
          5 &   &   & 15  &   &   \\
          6 & 7 &   & 16  &17   &   \\
            & 8 & 9 &   & 18  &19   \\
          \end{array}
          \right]
      \]
\end{frame}

\begin{frame}[fragile]
  \frametitle{Improvement: Block Sparse}
  Assume that B is decomposed into blocks which are either empty or `full'. `Full' blocks all have the same sparsity pattern. This is a sparse generalization of the block-sparse-column format.

  \begin{itemize}
  \item[$+$] We can reuse a single microkernel, reducing the number of instructions
  \item[$+$] The zero blocks reduce the number of filled-in nonzeros.
  \item[$-$] We can no longer visit blocks by using a loop with pointer arithmetic. \\Our options are:
    \begin{itemize}
    \item Look up the block location from a table in memory
    \item Unroll the block locations into the instruction stream and repeatedly jump to the microkernel.
    \end{itemize}
  \end{itemize}

      \[
      \left[
          \begin{array}{c c c | c c c }
          0 &   &   &    &    &    \\
          1 & 2 &   &    &    &    \\
            & 3 & 4 &    &    &    \\
          \hline
          5 &   &   & 10 &    &    \\
          6 & 7 &   & 11 & 12 &    \\
            & 8 & 9 &    & 13 & 14 \\
          \end{array}
          \right]
      \]
\end{frame}

\begin{frame}
  \frametitle{A family of dense x sparse algorithms}
  \include{flowchart}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Implementation}

  \begin{block}{Code Generation}
    \begin{itemize}
    \item Python domain-specific language for directly generating assembly
    \item Includes rich assembly AST: \texttt{Operand} $\leftarrow$ \texttt{Statement} $\leftarrow$ \texttt{Block} 
    \item Higher-level forms such as loops, subroutines, and jump tables
          implemented as plain Python functions which return a \texttt{Block}
    \item Pretty-printing, analysis, simulation are implemented using Visitor Pattern
    \item Interesting theoretical problem: Moving information from runtime to compile time in a general way
    \end{itemize}
  \end{block}

  \begin{block}{Matrix Cursor}
    A high-level abstraction for traversing and examining a sparse blocked matrix using \emph{logical coordinates}, automatically translated to physical memory offsets, for a given sparsity pattern and format. 
    \begin{itemize}
    \item \verb|has_nonzero_block(pointer, block_coords) -> bool|
    \item \verb|has_nonzero_cell(pointer, cell_coords) -> bool|
    \item \verb|move(pointer, block_coords) -> (new_pointer, statement)|
    \item \verb|look(pointer, cell_coords) -> (new_pointer, memory_address)|
    \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[plain,c]
\begin{center}
    \huge Thank you!
\end{center}
\end{frame}

\end{document}

### 2017-12-08 src/WavePropagation.cpp
Replaced u = T[ncells] with q = Q[ncells] = {T u, T hu}

@@ -47,10 +47,10 @@ T WavePropagation::computeNumericalFluxes(T dt)
 		T maxEdgeSpeed;
 
 		// Compute net updates
-		LaxFriedrichsFlux(m_u[i-1], m_u[i],
-				  dt, m_cellSize,
-				  m_uNetUpdatesLeft[i-1], m_uNetUpdatesRight[i-1],
-			          maxEdgeSpeed );
+		LaxFriedrichsFlux(m_q[i-1], m_q[i],
+				dt, m_cellSize,
+				m_uNetUpdatesLeft[i-1], m_uNetUpdatesRight[i-1],
+			    maxEdgeSpeed );
 
 		// Update maxWaveSpeed
 		if (maxEdgeSpeed > maxWaveSpeed){
@@ -68,34 +68,44 @@ void WavePropagation::updateUnknowns(T dt)
 {
   // Loop over all inner cells see Leveque p229 eq. (12.5)
    for (unsigned int i = 1; i < m_size+1; i++) {
-     m_u[i] -=  dt/m_cellSize * (m_uNetUpdatesRight[i-1] + m_uNetUpdatesLeft[i-1]);
+     m_q[i].h -=  dt/m_cellSize * (m_uNetUpdatesRight[i-1].h + m_uNetUpdatesLeft[i-1].h);
+     m_q[i].hu -=  dt/m_cellSize * (m_uNetUpdatesRight[i-1].hu + m_uNetUpdatesLeft[i-1].hu);
   }
 }
 
 void WavePropagation::setOutflowBoundaryConditions()
 {
-	m_u[0] = m_u[1]; m_u[m_size+1] = m_u[m_size];
+	m_q[0] = m_q[1]; 
+	m_q[m_size+1] = m_q[m_size];
 }
 
 void WavePropagation::setPeriodicBoundaryConditions()
 {
-  m_u[m_size+1] = m_u[1];
-  m_u[0] = m_u[m_size];
+  m_q[m_size+1] = m_q[1];
+  m_q[0] = m_q[m_size];
 }
 
 
 
 // See Leveque p. 234 eq 12.15
-void WavePropagation::LaxFriedrichsFlux(T u_l,T u_r,T dt, T dx, T& uNetUpdatesLeft, T& uNetUpdatesRight, T& maxEdgeSpeed)
[…]

### 2017-12-08 src/WavePropagation.h

@@ -73,28 +73,27 @@
 class WavePropagation
 {
 private:
-	T *m_u;
-
-	T *m_uNetUpdatesLeft;
-	T *m_uNetUpdatesRight;
-
+	Q *m_q;
+	// TODO: I assume these are supposed to be Q as well
+	Q *m_uNetUpdatesLeft;
+	Q *m_uNetUpdatesRight;
 	unsigned int m_size;
-
 	T m_cellSize;
+
 
 public:
 	/**
 	 * @param size Domain size (= number of cells) without ghost cells
 	 * @param cellSize Size of one cell
 	 */
-	WavePropagation(T* u,  unsigned int size, T cellSize)
-		: m_u(u),
+	WavePropagation(Q* q, unsigned int size, T cellSize)
+		: m_q(q),
 		  m_size(size),
 		  m_cellSize(cellSize)
 	{
 		// Allocate net updates
-		m_uNetUpdatesLeft = new T[size+1];
-		m_uNetUpdatesRight = new T[size+1];
+		m_uNetUpdatesLeft = new Q[size+1];
+		m_uNetUpdatesRight = new Q[size+1];
 	}
 
 	~WavePropagation()
@@ -135,7 +134,8 @@ public:
 	 * Compute fluctuations between left and right cell
 	 */
 
-	void LaxFriedrichsFlux(T u_l, T u_r,T dt, T dx, T& uNetUpdatesLeft, T& uNetUpdatesRight,T& maxEdgeSpeed);
+	void LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
+                         Q& uNetUpdatesLeft, Q& uNetUpdatesRight, T& maxEdgeSpeed);
 
 	/**
 	 * Flux function for the advection equation

### 2017-12-08 src/main.cpp

@@ -51,15 +51,12 @@ int main(int argc, char** argv)
 	// Scenario
 	scenarios::Gaussian scenario(args.size());
 
-	// Allocate memory
-	// Water height
-	T *u = new T[args.size()+2];
-	// Momentum
-
+	// Allocate memory across spatial domain
+	Q *q = new Q[args.size()+2];
 
 	// Initialize water height and momentum
-	for (unsigned int i = 0; i < args.size()+2; i++)
-	  u[i] = scenario.getU(i);
+	for (unsigned int x = 0; x < args.size()+2; x++)
+	  q[x] = scenario.getQ(x);
 
 
 	// Create a writer that is responsible printing out values
@@ -67,7 +64,7 @@ int main(int argc, char** argv)
 	writer::VtkWriter writer("adv1d", scenario.getCellSize());
 
 	// Helper class computing the wave propagation
-	WavePropagation wavePropagation(u, args.size(), scenario.getCellSize());
+	WavePropagation wavePropagation(q, args.size(), scenario.getCellSize());
 
 	// Write initial data
 	tools::Logger::logger.info("Initial data");
@@ -75,7 +72,7 @@ int main(int argc, char** argv)
 	// Current time of simulation
 	T t = 0;
 
-	writer.write(t, u, args.size());
+	writer.write(t, q, args.size());
 	
 	T maxTimeStep = args.size()/WavePropagation::ADVECTION_A * .4f;
 
@@ -98,11 +95,11 @@ int main(int argc, char** argv)
 		t += maxTimeStep;
 
 		// Write new values
-		writer.write(t, u, args.size());
+		writer.write(t, q, args.size());
 	}
 
 	// Free allocated memory
-	delete [] u;
+	delete [] q;
 
 	return 0;
 }

### 2017-12-08 src/scenarios/gaussian.h

@@ -59,14 +59,12 @@ public:
 	/**
 	 * @return Initial u at pos
 	 */
-	T getU(unsigned int pos)
+	Q getQ(unsigned int pos)
 	{
 	  T mid = m_size/2.0;
 	  T pos_rel = (mid-pos)/(m_size/10.0);
-	  return exp(-(pos_rel)*(pos_rel));
-	  
-	  
-	  //	  return 1.0;
+	  T h = exp(-(pos_rel)*(pos_rel));
+	  return {h, 0};
 	}
 
 	/**

### 2017-12-08 src/types.h

@@ -37,9 +37,17 @@
 #ifndef TYPES_H_
 #define TYPES_H_
 
-
 // Switch precision here
 typedef float T;
 //typedef double T;
+
+struct Q {
+	T h;
+	T hu;
+
+  //	Q operator-(Q & lhs, Q & rhs) {
+	//	return {lhs.h - rhs.h, lhs.hu - rhs.hu};
+	//}
+};
 
 #endif /* TYPES_H_ */

### 2017-12-08 src/writer/VtkWriter.h

@@ -102,7 +102,7 @@ public:
 	 *
 	 * @param size Number of cells (without boundary values)
 	 */
-	void write(const T time, const T *u,  unsigned int size)
+	void write(const T time, const Q *q,  unsigned int size)
 	{
 		// generate vtk file name
 		std::string l_fileName = generateFileName();
@@ -148,11 +148,16 @@ public:
 		vtkFile << "<CellData>" << std::endl;
 
 		// water surface height
-		vtkFile << "<DataArray Name=\"u\" type=\"Float32\" format=\"ascii\">" << std::endl;
+		vtkFile << "<DataArray Name=\"h\" type=\"Float32\" format=\"ascii\">" << std::endl;
 		for (int i=1; i < size+1; i++)
-				vtkFile << u[i] << std::endl;
+				vtkFile << q[i].h << std::endl;
 		vtkFile << "</DataArray>" << std::endl;
 
+		// water momentum
+		vtkFile << "<DataArray Name=\"hu\" type=\"Float32\" format=\"ascii\">" << std::endl;
+		for (int i=1; i < size+1; i++)
+      vtkFile << q[i].hu << std::endl;
+		vtkFile << "</DataArray>" << std::endl;
 
 		vtkFile << "</CellData>" << std::endl
 				<< "</Piece>" << std::endl;

### 2017-12-08 src/WavePropagation.cpp
Sanity-checking that advection still works

@@ -47,10 +47,9 @@ T WavePropagation::computeNumericalFluxes(T dt)
 		T maxEdgeSpeed;
 
 		// Compute net updates
-		LaxFriedrichsFlux(m_q[i-1], m_q[i],
-				dt, m_cellSize,
-				m_uNetUpdatesLeft[i-1], m_uNetUpdatesRight[i-1],
-			    maxEdgeSpeed );
+		LaxFriedrichsFlux(m_q[i-1], m_q[i], dt, m_cellSize,
+                      m_uNetUpdatesLeft[i-1], m_uNetUpdatesRight[i-1],
+                      maxEdgeSpeed );
 
 		// Update maxWaveSpeed
 		if (maxEdgeSpeed > maxWaveSpeed){
@@ -88,10 +87,10 @@ void WavePropagation::setPeriodicBoundaryConditions()
 
 
 // See Leveque p. 234 eq 12.15
-void WavePropagation::LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx, 
-	Q& uNetUpdatesLeft, 
-	Q& uNetUpdatesRight, 
-	T& maxEdgeSpeed)
+void WavePropagation::LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
+                                        Q& uNetUpdatesLeft,
+                                        Q& uNetUpdatesRight,
+                                        T& maxEdgeSpeed)
 {
 
   T flux_lh  = q_l.h  * ADVECTION_A;

### 2017-12-08 src/tools/args.h

@@ -62,7 +62,7 @@ private:
 public:
 	Args(int argc, char** argv)
 		: m_size(100),
-		  m_timeSteps(20.0)
+		  m_timeSteps(100.0)
 	{
 		const struct option longOptions[] = {
 			{"size", required_argument, 0, 's'},

### 2017-12-08 src/WavePropagation.cpp
Separated the pointwise flux calculation from the Lax-Friedrich; added SWE flux

@@ -84,7 +84,21 @@ void WavePropagation::setPeriodicBoundaryConditions()
   m_q[0] = m_q[m_size];
 }
 
+// Returns pointwise flux for advection. Updates maxEdgeSpeed reference. TODO: Do we want this?
+Q f_advection(Q q, T& maxEdgeSpeed) {
+  const T ADVECTION_A = 0.25;
+  T dh = q.h * ADVECTION_A;
+  T dhu = q.hu * ADVECTION_A;
+  maxEdgeSpeed = fabs(ADVECTION_A);
+  return {dh, dhu};
+}
 
+Q f_shallowwater(Q q, T& maxEdgeSpeed) {
+  T dh = q.hu;
+  const T G = 9.81;
+  T dhu = q.hu*q.hu/q.h + 0.5*G*q.h*q.h;
+  return {dh, dhu};
+}
 
 // See Leveque p. 234 eq 12.15
 void WavePropagation::LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
@@ -93,18 +107,15 @@ void WavePropagation::LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
                                         T& maxEdgeSpeed)
 {
 
-  T flux_lh  = q_l.h  * ADVECTION_A;
-  T flux_lhu = q_l.hu * ADVECTION_A;
-  T flux_rh  = q_r.h  * ADVECTION_A;
-  T flux_rhu = q_r.hu * ADVECTION_A;
+  Q flux_l = f_shallowwater(q_l, maxEdgeSpeed);
+  Q flux_r = f_shallowwater(q_r, maxEdgeSpeed);
   T a = dx/dt;
 
-  maxEdgeSpeed = fabs(ADVECTION_A);
   //  std::cout << maxEdgeSpeed << std::endl;
 
-  uNetUpdatesRight.h  = 0.5*((flux_rh - flux_lh) - a*(q_r.h - q_l.h));
-  uNetUpdatesRight.hu = 0.5*((flux_rhu - flux_lhu) - a*(q_r.hu - q_l.hu));
-  uNetUpdatesLeft.h   = 0.5*((flux_rh - flux_lh) + a*(q_r.h - q_l.h));
-  uNetUpdatesLeft.hu  = 0.5*((flux_rhu - flux_lhu) + a*(q_r.hu - q_l.hu));
[…]

### 2017-12-11 src/WavePropagation.cpp
Fixed bug, swe runs finally

@@ -67,8 +67,8 @@ void WavePropagation::updateUnknowns(T dt)
 {
   // Loop over all inner cells see Leveque p229 eq. (12.5)
    for (unsigned int i = 1; i < m_size+1; i++) {
-     m_q[i].h -=  dt/m_cellSize * (m_uNetUpdatesRight[i-1].h + m_uNetUpdatesLeft[i-1].h);
-     m_q[i].hu -=  dt/m_cellSize * (m_uNetUpdatesRight[i-1].hu + m_uNetUpdatesLeft[i-1].hu);
+     m_q[i].h -=  dt/m_cellSize * (m_uNetUpdatesRight[i].h + m_uNetUpdatesLeft[i-1].h);
+     m_q[i].hu -=  dt/m_cellSize * (m_uNetUpdatesRight[i].hu + m_uNetUpdatesLeft[i-1].hu);
   }
 }
 
@@ -94,9 +94,10 @@ Q f_advection(Q q, T& maxEdgeSpeed) {
 }
 
 Q f_shallowwater(Q q, T& maxEdgeSpeed) {
+  const T G = 1;  // To be compatible with picture on Ch 13, p. 257
   T dh = q.hu;
-  const T G = 9.81;
   T dhu = q.hu*q.hu/q.h + 0.5*G*q.h*q.h;
+  maxEdgeSpeed = std::max(q.hu/q.h - sqrt(G*q.h), q.hu/q.h + sqrt(G*q.h));
   return {dh, dhu};
 }
 
@@ -111,7 +112,7 @@ void WavePropagation::LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
   Q flux_r = f_shallowwater(q_r, maxEdgeSpeed);
   T a = dx/dt;
 
-  //  std::cout << maxEdgeSpeed << std::endl;
+  //std::cout << maxEdgeSpeed << std::endl;
 
   uNetUpdatesRight.h  = 0.5*((flux_r.h - flux_l.h) - a*(q_r.h - q_l.h));
   uNetUpdatesRight.hu = 0.5*((flux_r.hu - flux_l.hu) - a*(q_r.hu - q_l.hu));

### 2017-12-11 src/scenarios/gaussian.h

@@ -63,7 +63,7 @@ public:
 	{
 	  T mid = m_size/2.0;
 	  T pos_rel = (mid-pos)/(m_size/10.0);
-	  T h = exp(-(pos_rel)*(pos_rel));
+	  T h = 0.5 * exp(-(pos_rel)*(pos_rel)) + 1;
 	  return {h, 0};
 	}
 

### 2017-12-11 src/tools/args.h

@@ -62,7 +62,7 @@ private:
 public:
 	Args(int argc, char** argv)
 		: m_size(100),
-		  m_timeSteps(100.0)
+		  m_timeSteps(20.0)
 	{
 		const struct option longOptions[] = {
 			{"size", required_argument, 0, 's'},

### 2017-12-11 src/WavePropagation.cpp
Candidate Rusanov. Also now plotting maxEdgeSpeed.

@@ -48,10 +48,11 @@ T WavePropagation::computeNumericalFluxes(T dt)
 
 		// Compute net updates
 		LaxFriedrichsFlux(m_q[i-1], m_q[i], dt, m_cellSize,
-                      m_uNetUpdatesLeft[i-1], m_uNetUpdatesRight[i-1],
+                      m_updatesLeft[i-1], m_updatesRight[i-1],
                       maxEdgeSpeed );
 
 		// Update maxWaveSpeed
+    m_maxEdgeSpeed[i] = maxEdgeSpeed;
 		if (maxEdgeSpeed > maxWaveSpeed){
 		  maxWaveSpeed = maxEdgeSpeed;
 		}
@@ -67,8 +68,8 @@ void WavePropagation::updateUnknowns(T dt)
 {
   // Loop over all inner cells see Leveque p229 eq. (12.5)
    for (unsigned int i = 1; i < m_size+1; i++) {
-     m_q[i].h -=  dt/m_cellSize * (m_uNetUpdatesRight[i].h + m_uNetUpdatesLeft[i-1].h);
-     m_q[i].hu -=  dt/m_cellSize * (m_uNetUpdatesRight[i].hu + m_uNetUpdatesLeft[i-1].hu);
+     m_q[i].h -=  dt/m_cellSize * (m_updatesRight[i].h + m_updatesLeft[i-1].h);
+     m_q[i].hu -=  dt/m_cellSize * (m_updatesRight[i].hu + m_updatesLeft[i-1].hu);
   }
 }
 
@@ -97,26 +98,38 @@ Q f_shallowwater(Q q, T& maxEdgeSpeed) {
   const T G = 1;  // To be compatible with picture on Ch 13, p. 257
   T dh = q.hu;
   T dhu = q.hu*q.hu/q.h + 0.5*G*q.h*q.h;
-  maxEdgeSpeed = std::max(q.hu/q.h - sqrt(G*q.h), q.hu/q.h + sqrt(G*q.h));
+  maxEdgeSpeed = std::max(fabs(q.hu/q.h - sqrt(G*q.h)),
+                          fabs(q.hu/q.h + sqrt(G*q.h)));
   return {dh, dhu};
 }
 
 // See Leveque p. 234 eq 12.15
 void WavePropagation::LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
-                                        Q& uNetUpdatesLeft,
[…]

### 2017-12-11 src/WavePropagation.h

@@ -72,16 +72,15 @@
  */
 class WavePropagation
 {
-private:
+public:
 	Q *m_q;
-	// TODO: I assume these are supposed to be Q as well
-	Q *m_uNetUpdatesLeft;
-	Q *m_uNetUpdatesRight;
+	Q *m_updatesLeft;
+	Q *m_updatesRight;
+	T *m_maxEdgeSpeed;
 	unsigned int m_size;
 	T m_cellSize;
 
 
-public:
 	/**
 	 * @param size Domain size (= number of cells) without ghost cells
 	 * @param cellSize Size of one cell
@@ -92,15 +91,17 @@ public:
 		  m_cellSize(cellSize)
 	{
 		// Allocate net updates
-		m_uNetUpdatesLeft = new Q[size+1];
-		m_uNetUpdatesRight = new Q[size+1];
+		m_updatesLeft = new Q[size+1];
+		m_updatesRight = new Q[size+1];
+		m_maxEdgeSpeed = new T[size+1];
 	}
 
 	~WavePropagation()
 	{
 		// Free allocated memory
-		delete [] m_uNetUpdatesLeft;
-		delete [] m_uNetUpdatesRight;
+		delete [] m_updatesLeft;
+		delete [] m_updatesRight;
+    delete [] m_maxEdgeSpeed;
 	}
 
 	/**
@@ -135,7 +136,7 @@ public:
 	 */
 
 	void LaxFriedrichsFlux(Q q_l, Q q_r, T dt, T dx,
-                         Q& uNetUpdatesLeft, Q& uNetUpdatesRight, T& maxEdgeSpeed);
+                         Q& updatesLeft, Q& updatesRight, T& maxEdgeSpeed);
 
 	/**
 	 * Flux function for the advection equation

### 2017-12-11 src/main.cpp

@@ -72,7 +72,7 @@ int main(int argc, char** argv)
 	// Current time of simulation
 	T t = 0;
 
-	writer.write(t, q, args.size());
+	writer.write(t, q, wavePropagation.m_maxEdgeSpeed, args.size());
 	
 	T maxTimeStep = args.size()/WavePropagation::ADVECTION_A * .4f;
 
@@ -95,7 +95,7 @@ int main(int argc, char** argv)
 		t += maxTimeStep;
 
 		// Write new values
-		writer.write(t, q, args.size());
+		writer.write(t, q, wavePropagation.m_maxEdgeSpeed, args.size());
 	}
 
 	// Free allocated memory

### 2017-12-11 src/writer/VtkWriter.h

@@ -102,7 +102,7 @@ public:
 	 *
 	 * @param size Number of cells (without boundary values)
 	 */
-	void write(const T time, const Q *q,  unsigned int size)
+	void write(const T time, const Q *q, const T *maxEdgeSpeed, unsigned int size)
 	{
 		// generate vtk file name
 		std::string l_fileName = generateFileName();
@@ -158,6 +158,12 @@ public:
 		for (int i=1; i < size+1; i++)
       vtkFile << q[i].hu << std::endl;
 		vtkFile << "</DataArray>" << std::endl;
+
+		// water momentum
+		vtkFile << "<DataArray Name=\"maxEdgeSpeed\" type=\"Float32\" format=\"ascii\">" << std::endl;
+		for (int i=1; i < size+1; i++)
+      vtkFile << maxEdgeSpeed[i] << std::endl;
+    vtkFile << "</DataArray>" << std::endl;
 
 		vtkFile << "</CellData>" << std::endl
 				<< "</Piece>" << std::endl;

### 2017-12-11 src/main.cpp
Breaking dam scenario. Breaks older version of Paraview (but not 5.4) on timestep 45, not sure why.

@@ -38,6 +38,7 @@
 #include "types.h"
 #include "WavePropagation.h"
 #include "scenarios/gaussian.h"
+#include "scenarios/breakingdam.h"
 #include "writer/VtkWriter.h"
 #include "tools/args.h"
 
@@ -49,7 +50,8 @@ int main(int argc, char** argv)
 	tools::Args args(argc, argv);
 
 	// Scenario
-	scenarios::Gaussian scenario(args.size());
+	scenarios::BreakingDam scenario(args.size());
+	//scenarios::Gaussian scenario(args.size());
 
 	// Allocate memory across spatial domain
 	Q *q = new Q[args.size()+2];
@@ -82,8 +84,8 @@ int main(int argc, char** argv)
 				<< " at time " << t << std::endl;
 
 		// Update boundaries
-		//		wavePropagation.setOutflowBoundaryConditions();
-		wavePropagation.setPeriodicBoundaryConditions();		
+		wavePropagation.setOutflowBoundaryConditions();
+		//wavePropagation.setPeriodicBoundaryConditions();		
 
 		// Compute numerical flux on each edge
 		maxTimeStep = wavePropagation.computeNumericalFluxes(maxTimeStep);

### 2017-12-11 src/scenarios/breakingdam.h

@@ -1,0 +1,86 @@
+/**
+ * @file
+ *  This file is part of SWE1D
+ *
+ *  SWE1D is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  SWE1D is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with SWE1D.  If not, see <http://www.gnu.org/licenses/>.
+ *
+ *  Diese Datei ist Teil von SWE1D.
+ *
+ *  SWE1D ist Freie Software: Sie koennen es unter den Bedingungen
+ *  der GNU General Public License, wie von der Free Software Foundation,
+ *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
+ *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
+ *
+ *  SWE1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
+ *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
+ *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
+ *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
+ *
+ *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
+ *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
+ * 
+ * @copyright 2013 Technische Universitaet Muenchen
+ * @author Sebastian Rettenberger <rettenbs@in.tum.de>
+ */
+
+#ifndef SCENARIOS_BREAKINGDAM_H_
+#define SCENARIOS_BREAKINGDAM_H_
+
+#include "types.h"
+#include <iostream>
+#include <math.h>
+
+namespace scenarios
+{
+
+class BreakingDam
+{
+private:
+	/** Number of cells */
[…]

### 2017-12-20 paper/paper.tex
Added latex template

\input{paper-layout}
\input{german}
\pagestyle{empty}  % Skip page number on first page

\hyphenation{}

\begin{document}

\title{Solving the Shallow Water Equations\\ using Finite Volumes and Lax-Friedrichs}

\specialpapernotice{Fundamentals of Wave Simulation Seminar}

\author{
\authorblockN{Nathan W Brei}
\authorblockA{Fakultät für Informatik\\Technische Universität München\\
Email: brei@in.tum.de} 
%\and
%\authorblockN{}
%\authorblockA{}
}

\maketitle

% Re-enable page number on first page
%\thispagestyle{plain}

\input{abstract}

\begin{keywords}
	Shallow water equations, Finite volumes, Lax-Friedrichs method, Rusanov, artificial viscosity, Julia
\end{keywords}

\section{Introduction}

This paper has three aims. Firstly, it shall introduce the shallow water equations and describe how finite volume methods can be used to compute a numerical solution. Secondly, it shall explore two such methods, the Lax-Friedrich and the Local-Lax-Friedrich (Rusanov) methods, describing them conceptually via equations and concretely via code snippets. The code is organized to map to the concepts as cleanly as possible and to leverage higher-order programming techniques and patterns. Finally, the code shall be used to generate solutions to several model problems, which can be examined in order to gain insight into the concept and consequences of numerical viscosity.

\section{Shallow Water Equations}

The shallow water equations are a system of hyperbolic partial differential equations which are effective at describing the behavior of tsunamis and overland flooding. They are derived from the general Navier-Stokes equations by making the following assumptions:

\begin{itemize}
\item In the 1-D case, the flow is in a channel of unit width
\item Horizontal velocity $u(x)$ across any cross section $x$ is constant
\item Vertical velocity $v(x)$ is small (but not zero) and falls out of the equations
\item Pressure is dominated by hydrostatics: $p=\frac{1}{2}\rho gh^2$
\end{itemize}

The 1D equations are as follows:

\[
\begin{bmatrix}h \\ hu \end{bmatrix}_t + \begin{bmatrix}hu \\ hu^2 + \frac{1}{2}gh^2\end{bmatrix}_x = 0 \]

This paper chooses the state variables $q := (h, hu)$ to form a state space, $Q$. This can be expressed in code as follows:
\begin{verbatim}
const Float = Float64

mutable struct Q
    h :: Float
    hu :: Float
end
\end{verbatim}

There are several tradeoffs associated with this representation of state space. In performance-critical code, the choice of data structure is driven primarily by memory access patterns, which are not considered here. The principal benefit of this representation is that it moves some of the semantics of the problem domain into the type system, which would permit future extensions such as awareness of physical units. On the other hand, this representation is unaware of the fact that $Q$ forms a vector space, resulting in duplicated, messier, and possibly less efficient code.

This latter problem may be resolved quickly, as Julia uses \emph{multimethods}. Multimethods

\begin{verbatim}
import Base.+
import Base.-
import Base.*

function +(lhs::Q, rhs::Q)
    Q(lhs.h + rhs.h, lhs.hu + rhs.hu)
end

function -(lhs::Q, rhs::Q)
    Q(lhs.h - rhs.h, lhs.hu - rhs.hu)
end

function *(a::Real, q::Q)
    Q(a*q.h, a*q.hu)
end
\end{verbatim}

The shallow water equations may be defined over this state space in an abstract way. Because they match the pattern of a general conservation law,

\[q_t(x,t) + f(q(x,t))_x = 0\]

their dynamics may be captured as a flux function, $f$, such that

\[f : (q_1, q_2) \mapsto \begin{bmatrix} q_2\\ q_2^2/q_1 + \frac{1}{2}gq_1^2\end{bmatrix}\]

This flux function may be expressed in Julia directly:

\begin{verbatim}
const G = 9.81f0
function f(q :: Q)
    Q(q.hu, q.hu^2/q.h + 0.5*G*q.h^2)
end
\end{verbatim}

The shallow water equations enter the numerical code in only one other place, the max wave speed function. The physical wave speeds are given by the eigenvalues of the flux Jacobian

\[f'(q) = \begin{bmatrix}0 & 1 \\ -u^2 + gh & 2u \end{bmatrix}\]

\[\lambda_{max} = \max\ \lvert u \pm \sqrt{gh} \rvert \]

\begin{verbatim}
function wavespeed(q :: Q)
    u = q.hu / q.h
    c = sqrt(G * q.h)
    max(abs(u-c), abs(u+c))
end
\end{verbatim}

\subsection{Unterkapitel}

blabla mit drei Quellenangaben\cite{ietf-ipfix-protocol,snoeren2001hash,belenky2003ip}

\begin{figure}[h]%
 	\begin{center}%
 		\includegraphics[scale=0.1]{figure1.png}%
 		\caption{Baum}\label{fig:baum}%
 	\end{center}%
\end{figure}

\begin{table}[h]%
 	\begin{center}%
		\caption{Beispieltabelle}\label{tab:example}%
	 	\begin{tabular}{c|c}%
 			Spalte1 & Spalte2\\
 			\hline
 			0 & 1\\
 		\end{tabular}%
 	\end{center}%
\end{table}

%\input{section2}

\input{conclusion}

%\section*{Acknowledgment}
%\addcontentsline{toc}{section}{Acknowledgment}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section
% NOTE: BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/

% can use a bibliography generated by BibTeX as a .bbl file
% standard IEEE bibliography style from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/bibtex
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,references}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{ref:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to {\LaTeX}}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

\end{document}

### 2017-12-21 paper/paper.tex
Getting closer

@@
 \input{german}
+\usepackage{amsmath}
 \pagestyle{empty}  % Skip page number on first page
@@
 
-This latter problem may be resolved quickly, as Julia uses \emph{multimethods}. Multimethods
+This latter problem may be resolved quickly, thanks to Julia's type system.
 
@@
 
-The shallow water equations enter the numerical code in only one other place, the max wave speed function. The physical wave speeds are given by the eigenvalues of the flux Jacobian
+The shallow water equations enter the numerical code in only one other place, the max wave speed function. The physical speeds of wave propagation are given by the eigenvalues of the flux Jacobian,
 
@@
 
+and the absolute value of the dominating eigenvalue,
+
 \[\lambda_{max} = \max\ \lvert u \pm \sqrt{gh} \rvert \]
+
+will be shown to play an important role in ensuring numerical stability. In Julia it is calculated as:
 
@@
 
-\subsection{Unterkapitel}
+\section{Finite Volumes}
 
-blabla mit drei Quellenangaben\cite{ietf-ipfix-protocol,snoeren2001hash,belenky2003ip}
+Finite volumes are a family of numerical methods which are well suited for computing an approximate solution to hyperbolic systems such as the shallow water equations.
+They have two main advantages over other methods, such as finite differences and finite elements.
+Firstly, they automatically conserve physical quantities $q$.
+Secondly, they can capture discontinuities, such as shocks, which is considerably simpler than explicitly tracking them.
+The main downside is that the profile of the shock will be smeared by the spatial discretization.
+
+The starting point is to recast the conservation equations in integral form:
+
+\[\frac{d}{dt} \int_{C_i} q(x,t) dx = f(q(x_{i-1/2},t)) - f(q(x_{1+1/2},t))\]
+
+Since these equations hold for an arbitrary control volume, it makes sense to partition the spatial domain into cells.
+For simplicity these cells are assumed to be regular, although the equations hold just as well over an unstructured mesh.
+
+\[C_i := (x_{i-1/2}, x_{i+1/2})\]
+
+After integrating with respect to time, each term takes on a physical significance directly relevant to our model:
+
+\begin{multline}
+\int_{C_i} q(x,t_{n+1})dx - \int_{C_i} q(x,t_n) dx = \\ \int_{t_n}^{t_{n+1}} f(q(x_{i-1/2},t))dt - \int_{t_n}^{t_{n+1}} f(q(x_{i+1/2},t))dt
+\end{multline}
+
+\begin{itemize}
+\item Let $Q_{i}^{n} \approx \frac{1}{\Delta x} \int_{C_i} q(x,t_n) dx $ be the approximate average $q$ in cell $i$ at time $n$.
+
+\item Let $F_{i+1/2}^{n} \approx \frac{1}{\Delta t} \int_{t_n}^{t_{n+1}} f(q(x_{i+1/2},t))dt$ be the approximate average flux $f$ from cell $i$ into $i+1$.
+\end{itemize}
+Rearranging leads to a general update scheme:
+
+\[Q_{i}^{n+1} = Q_{i}^{n} - \frac{\Delta t}{\Delta x}(F_{i+1/2}^{n} - F_{i-1/2}^{n}) \]
+
+Ideally, the flux function can be calculated using only the state in the adjacent cells, which implies a fully discrete method:
+
+\[F_{i+1/2}^{n} = \mathcal{F}(Q_{i}^n, Q_{i+1}^n)\]
+
+Taking a step back, it is clear that these also fall under the umbrella of finite difference methods.
+However, they have the additional property of being conservative: Even if the computed fluxes are rather inaccurate, the sum of fluxes over the spatial domain will vary only according to boundary conditions.
+
+The desired properties of finite volumes hold for any choice of $F_{i+1/2}^{n}$, as long as the resulting method is consistent and stable.
+Similarly, they hold for any choice of hyperbolic PDE, which is abstracted away as a state space $Q_i$, a pointwise flux $f(Q_i)$ and max wavespeed $\lambda_{max}(Q_i)$.
+Thus, the implementation may be organized into an abstract finite volume method which is parameterized with different strategies.
+
+\section{Code architecture}
+
+Let us now consider a clean organizing principle for the code.
+
+Thanks to parametric polymorphism, the spatial domain can be quickly defined as a type alias \verb!QS = Array{Q,1}!
+This states that spatial variables are implemented as a 1-dimensional array with inner type \verb!Q!
+as defined above.
+There is an additional constraint in that there is a ghost cell on either side, so the `true' spatial domain ranges over the indices $[2, n_{cells}+1]$.
+
+Two improvements are possible.
+Firstly, by using \verb!OffsetArray!s instead of regular \verb!Array!s, the indexes can be cheaply translated to $[1, n_{cells}]$.
+Secondly, the type QS can be parameterized by the number of cells, \verb!QS{ncells}!, which would ensure that compatibly sized arrays are always used, that ghost cells always exist, and that the array sizes are known at compile time, allowing for additional compiler optimizations.
+Though potentially useful for larger projects, these features would introduce more complexity than they conceal.
+
+\begin{itemize}
+	\item Flux function f
+	\item Initial conditions
+	\item Boundary conditions
+	\item Timestep size
+\end{itemize}
+
+In traditional object-oriented programming, this decoupling is usually achieved via the Strategy Pattern. Because Julia has first-class functions (lambdas), it can be achieved much more tersely.
+
+\section{Numerical Stability}
+
+In order to ensure numerical stability, a necessary (but not sufficient) condition that must be met is the \emph{CFL condition}.
+For a hyperbolic system, information propagates through the spatial domain at finite speeds.
+Since the system evolves by exchanging fluxes between neighboring cells, it is intuitively necessary that the timestep size be small enough that the fluxes don't travel further than one cell per timestep.
+More precisely, the \emph{numerical domain of dependence} (the set of cells whose old values factor in to the new value of each cell) must contain the \emph{physical domain of dependence}, the region of space from which waves propagated.
+This is captured as a relation between cell size, timestep size, and wave speed, called the Courant number:
+
+\[ \nu := \Bigg\lvert\frac{\bar{u}\Delta t}{\Delta x} \Bigg\rvert = \frac{\Delta t}{\Delta x} \lvert \lambda_{max} \rvert  \leq 1 \]
+
+Because this code uses the same timestep size for all cells, the maximum wave speed must be calculated across the entire spatial domain for each timestep iteration.
+The maximum wavespeed calculated from the flux Jacobian is exact only in the small-perturbation case, so an additional correction factor, $\mu$, scales the timestep down.
+Thus the experimenter has two knobs for balancing speed, accuracy, and stability: The spatial discretization $\Delta x$ and the timestep correction $\mu$.
+
+In Julia, the CFL condition is encoded as follows. One useful feature is the function chain operator \verb!|>!, which aids in readability.
+
+\begin{verbatim}
+function cfl_dt(qs::QS, dx::Float, mu::Float)
+    lambda = map(wavespeed,qs) |> maximum
+    dx / lambda * mu
+end
+\end{verbatim}
+
+\section{Model Problems}
+
+\section{The Lax-Friedrichs Method}
+
+Having built up a skeleton of a finite volume method, it is finally time to furnish the most essential missing piece, the cell-wise flux function.
+The most obvious choice of $F$ is simply the average of the fluxes at the center of either cell.
+
+\[ F_{i+1/2} = \frac{1}{2}\big[f(Q_{i} + f(Q_{i+1})\big] \]
+
+\[Q_i^{n+1} = Q_i^n - \frac{\Delta t}{2 \Delta x}\big[f(Q_{i+1}^n - f(Q_{i-1}^n))] \]
+
+This yields a forward-time, centered-space finite difference scheme.
+Unfortunately a von Neumann analysis shows that it is always unstable.
+The Lax-Friedrichs method simply modifies this scheme by changing the time difference to use a value of $Q_i^n$ averaged from the neighboring cells:
+
+\[Q_i^n \approx \frac{1}{2}(Q_{i-1}^n + Q_{i+1}^n)\]
+
+The resulting update scheme becomes:
+
+\[ Q_i^{n+1} = \frac{1}{2}\Big[Q_{i-1}^n + Q_{i+1}^n\Big] - \frac{\Delta t}{2 \Delta x} \Big[f(Q_{i+1}^n) - f(Q_{i-1}^n)\Big]\]
+
+which may be rearranged to yield a cell-wise flux function:
+
+\[ F_{i+1/2} = \frac{1}{2} \Big[ f(Q_{i+1}) + f(Q_{i}) - \frac{\Delta x}{\Delta t}(Q_{i+1} - Q_{i})\Big] \]
+
+This can be encoded as a function 
+* Input:  Buffers for $q(x), F_l(x), F_r(x)$, cell and timestep sizes
+* Output: Updated inter-cell flux buffers
+
+\begin{verbatim}
+function lxf!(qs::QS, Fl::QS, Fr::QS, 
+			  ncells::Int, 
+              dx::Float, dt::Float)
+
+    a = dx/dt
+    for x = 2:ncells+2
+        ql,qr = qs[x-1], qs[x]
+        fl,fr = f(ql), f(qr)
+
+        Fr[x-1] = 0.5*((fr-fl) - a*(qr - ql))
+        Fl[x-1] = 0.5*((fr-fl) + a*(qr - ql))
+    end
+end
+\end{verbatim}
+
+\section{Lax-Friedrichs Results}
+
+The Lax-Friedrichs method implemented was run against two model problems.
+The first used a Gaussian perturbation for its initial condition with reflecting boundary conditions.
+The second used a breaking dam initial condition with outflow boundary conditions.
+Two experiments were run in order to examine the behavior of the knobs $n_{cells}$  and $\mu$ independently.
+These are depicted in Figures~\ref{fig:lxf_gaussian_ncells},~\ref{fig:lxf_breakingdam_ncells}, and~\ref{fig:lxf_breakingdam_mu}.
 
@@
  	\begin{center}%
- 		\includegraphics[scale=0.1]{figure1.png}%
+ 		\includegraphics[scale=0.6]{../plots/lxf_gaussian_ncells.png}%
+ 		\caption{LxF performance relative to $n_{cells}$}\label{fig:lxf_gaussian_ncells}%
+ 	\end{center}%
+\end{figure}
+
+\begin{figure}[h]%
+ 	\begin{center}%
+ 		\includegraphics[scale=0.6]{../plots/lxf_breakingdam_ncells.png}%
+ 		\caption{LxF performance relative to $n_{cells}$}\label{fig:lxf_breakingdam_ncells}%
+ 	\end{center}%
+\end{figure}
+
+\begin{figure}[h]%
+ 	\begin{center}%
+ 		\includegraphics[scale=0.6]{../plots/lxf_breakingdam_mu.png}%
+ 		\caption{LxF performance relative to $\mu$}\label{fig:lxf_breakingdam_mu}%
+ 	\end{center}%
+\end{figure}
+
+The experiment depicted in Figure~\ref{fig:lxf_gaussian_ncells} is rather intuitive.
+A Gaussian perturbation propagates symmetrically across the spatial domain.
+As the spatial discretization is refined, a shock wave gradually emerges from each (previously symmetric) ripple.
+It should be noted that these shock waves are smeared across tens or hundreds of cells, far more than the finite volume method presumably requires.
+Furthermore, the crest is reduced and the troughs are filled in: Decreasing the spatial discretization appears to increase dissipation.
+
+Applying the same experiment to the breaking dam scenario, shown in Figure~\ref{fig:lxf_breakingdam_ncells} yields similar observations, with an unexpected twist.
+As the mesh becomes coarser, a nonphysical stairstep pattern emerges.
+This can be directly traced back to the update rule: $Q_i^{n+1}$ depends on $Q_{i+1}^n$ and $Q_{i-1}^n$, but not $Q_{i}^n$.
+As a result, the odd and even grid cells evolve independently from each other and any disparity between the two oscillates without dampening.
+Thus the stairstep pattern did not manifest in the Gaussian scenario because the initial condition was smooth.
+
+The final experiment considers the effect of the timestep correction $\mu$.
+It uses a fine mesh in order to reduce the dissipation described previously.
+As defined earlier, $\mu$ is a constant factor which scales down the timestep in order to maintain stability even in the presence of large nonlinearities.
+However, the behavior shown is remarkably counterintuitive: As the timestep is decreased, the solution becomes more dissipative, and consequently less accurate.
+When $\mu=1.0$, the shock is crisp and almost vertical; when $\mu=0.1$, the shock is effectively lost.
+The next section shall explore the cause of this phenomenon by showing how viscous behavior can arise as a result of the numerical method used, rather than the underlying physical system.
+
+\section{Artificial Viscosity}
+
+The key idea behind artificial viscosity is that an approximate solution to a certain problem is oftentimes an exact solution to an approximation of that problem.
+The trick is to regard simple discretizatons of a PDE as being building blocks which can be assembled into more sophisticated (and correspondingly more approximate) discretizations.
+One can then map the sophisticated discretization directly back to a modified PDE.
+In the case of Lax-Friedrichs, the approximation chosen for the flux function is effectively adding a diffusive term to the shallow-water equations.
+
+Suppose that the `pure' solution to the shallow water equations, $q_t + {f(q)}_x = 0$, corresponds to the naive flux function
+
+\[F_{i+1/2} = \frac{1}{2} \Big[f(Q_{i+1} + f(Q_{i})\Big]\]
+
+even though this is known to be unstable. Similarly, suppose that the `pure' solution to the diffusion equation, $q_t = \beta q_{xx}$, corresponds to the flux function
+
+\[F_{i+1/2} = -\beta \frac{Q_{i+1} - Q_{i}}{\Delta x}\]
+
+This can be derived by a straightforward finite difference approximation. The Lax-Friedrichs flux function shown previously can be trivially rearranged as:
+
+\[ F_{i+1/2} = \frac{1}{2} \Big[f(Q_{i}) + f(Q_{i+1})\Big] - \frac{\Delta x}{2\Delta t}\Big[Q_{i+1} - Q_{i}\Big] \]
+
+It is clear that this flux function pattern-matches the discretization of a different PDE:
+
+\[ q_t + f(q)_x = \frac{(\Delta x)^2}{2\Delta t} q_{xx} \]
+
+Thus the Lax-Friedrichs parameter $a = \frac{\Delta x}{\Delta t}$ controls the diffusion coefficient $\beta$ according to
+
+\[\beta = \frac{(\Delta x)^2}{2\Delta t} = \frac{1}{2} a \Delta x \]
+
+This explains the two results described previously.
+Firstly, by taking $a$ as constant, and reducing the cell size $\Delta x \to 0$, it is clear that $\beta \to 0$ and the viscous term in the PDE vanishes.
+Thus the Lax-Friedrichs discretization remains consistent with the pure shallow-water equations.
+
+Secondly, reducing the timestep size $\Delta t$ increases $a$, which then increases $\beta$.
+This explains the counterintuitive phenomenon where the solution becomes increasingly dissipative, and hence less accurate, as the timestep size is decreased.
+This then raises the question of whether it is possible to decouple $\beta$ from $\Delta t$ while still maintaining a consistent and stable discretization.
+The method described in the following section mostly achieves this.
+
+\section{The Local-Lax-Friedrichs Method}
+
+Local-Lax-Friedrichs is a modification which consciously uses artificial viscosity in a clever way.
+In the basic Lax-Friedrichs, the viscosity parameter $a$ is a constant derived from the timestep $\Delta t$, which in turn is derived from the CFL condition.
+Although the timestep is contrained to be constant over the entire spatial domain (local time stepping is not considered in the scope of this paper), the viscosity is not.
+On the contrary, varying the viscosity over the spatial domain is advantageous: in regions with shocks and other discontinuities, the presence of viscosity helps ensure numerical stability, and in smooth regions, the lack of viscosity leads to a less dissipative and hence more accurate solution.
+
+$a_{i-1/2} = \max(\lvert f'(q) \rvert) \ \forall q \in (Q_{i-1}, Q_{i})$
+
+Then $ F_{i-1/2} = \frac{1}{2}[f(Q_{i-1}) + f(Q_i) - a_{i-1/2}(Q_i - Q_{i-1})]$
+
+From the CFL condition, it is clear that this approach is strictly less viscous than basic Lax-Friedrichs:
+
+\[\lvert f'(q)\rvert \leq \frac{\Delta x}{\Delta t}\]
+
+\begin{figure}[h]%
+ 	\begin{center}%
+ 		\includegraphics[scale=0.6]{../plots/artificial_viscosity.png}%
+ 		\caption{Wavespeed over the spatial domain}\label{fig:artificial_viscosity}%
+ 	\end{center}%
+\end{figure}
+
+This is represented in code as follows. The flux
+
+\begin{verbatim}
+function llxf!(qs::QS, Fl::QS, Fr::QS, 
+              ncells::Int, 
+              dx::Float, dt::Float)
+
+    for x = 2:ncells+2
+        ql,qr = qs[x-1], qs[x]
+        fl,fr = f(ql), f(qr)
+
+        a = max(wavespeed(ql), wavespeed(qr))
+
+        Fr[x-1] = 0.5*((fr-fl) - a*(qr-ql))
+        Fl[x-1] = 0.5*((fr-fl) + a*(qr-ql))
+    end
+end
+\end{verbatim}
+
+\section{Comparison of results}
+\begin{figure}[h]%
+ 	\begin{center}%
+ 		\includegraphics[scale=0.6]{../plots/lxf_llxf_breakingdam.png}%
+ 		\caption{Baum}\label{fig:baum}%
+ 	\end{center}%
+\end{figure}
+\begin{figure}[h]%
+ 	\begin{center}%
+ 		\includegraphics[scale=0.6]{../plots/lxf_llxf_gaussian.png}%
  		\caption{Baum}\label{fig:baum}%
@@
 
-\begin{table}[h]%
- 	\begin{center}%
-		\caption{Beispieltabelle}\label{tab:example}%
-	 	\begin{tabular}{c|c}%
- 			Spalte1 & Spalte2\\
- 			\hline
- 			0 & 1\\
- 		\end{tabular}%
- 	\end{center}%
-\end{table}
+\begin{itemize}
+\item For Lax-Friedrich, $\mu$ affects dissipation
+\item For Local-Lax-Friedrich, $\mu$ affects dissipation much less
+\item Lax-Friedrich approaches Local-Lax-Friedrich as $\mu \to 1.0$
+\item Both methods are unstable for $\mu > 1.0$ due to CFL violation
+\end{itemize}
 
-%\input{section2}
+\section{Conclusion}
 
-\input{conclusion}
+%\input{conclusion}
 

### 2017-12-21 paper/paper.tex
Conclusion and bibliography

@@
 
-\input{abstract}
+%\input{abstract}
 
-\begin{keywords}
-	Shallow water equations, Finite volumes, Lax-Friedrichs method, Rusanov, artificial viscosity, Julia
-\end{keywords}
+%\begin{keywords}
+%	Shallow water equations, Finite volumes, Lax-Friedrichs method, Rusanov, artificial viscosity, Julia
+%\end{keywords}
 
@@
 
-The shallow water equations are a system of hyperbolic partial differential equations which are effective at describing the behavior of tsunamis and overland flooding.
-They are derived from the general Navier-Stokes equations by making the following assumptions:
+The shallow water equations are a system of hyperbolic partial differential equations which are effective at describing the behavior of tsunamis and overland flows.\cite{delestre:hal-00932234} They are derived from the general Navier-Stokes equations by making the following assumptions:
 
@@
 
-This latter problem may be resolved quickly, thanks to Julia's type system.
+This latter problem may be resolved quickly, thanks to Julia's type system.\cite{doi:10.1137/141000671}
 
@@
 
-Finite volumes are a family of numerical methods which are well suited for computing an approximate solution to hyperbolic systems such as the shallow water equations.
-They have two main advantages over other methods, such as finite differences and finite elements.
+Finite volumes are a family of numerical methods which are well suited for computing an approximate solution to hyperbolic systems such as the shallow water equations.\cite{leveque2002finite} They have two main advantages over other methods, such as finite differences and finite elements.
 Firstly, they automatically conserve physical quantities $q$.
@@
 
-Ideally, the flux function can be calculated using only the state in the adjacent cells, which implies a fully discrete method:
+Ideally, the flux function can be calculated using only the state in the adjacent cells, which yields a fully discrete method:
 
@@
 This yields a forward-time, centered-space finite difference scheme.
-Unfortunately a von Neumann analysis shows that it is always unstable.
-The Lax-Friedrichs method simply modifies this scheme by changing the time difference to use a value of $Q_i^n$ averaged from the neighboring cells:
+Unfortunately a von Neumann analysis shows that it is always unstable.\cite{strang2007computational} The Lax-Friedrichs method simply modifies this scheme by changing the time difference to use a value of $Q_i^n$ averaged from the neighboring cells:
 
@@
 
-\section{Comparison of results}
+The Local Lax-Friedrichs method does not entirely break the coupling between dissipation and $\mu$, but it strongly reduces it, as depicted in Figure~\ref{fig:llxf_breakingdam_mu}.
+This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, only for LLxF.
+\begin{figure}[h]%
+    \begin{center}%
+        \includegraphics[scale=0.6]{../plots/llxf_breakingdam_mu.png}%
+        \caption{Baum}\label{fig:llxf_breakingdam_mu}%
+    \end{center}%
+\end{figure}
+
+\section{LxF vs LLxF Comparison}
+
 \begin{figure}[h]%
@@
         \includegraphics[scale=0.6]{../plots/lxf_llxf_breakingdam.png}%
- 		\caption{Baum}\label{fig:baum}%
+        \caption{Baum}\label{fig:cmp_breakingdam}%
     \end{center}%
@@
         \includegraphics[scale=0.6]{../plots/lxf_llxf_gaussian.png}%
- 		\caption{Baum}\label{fig:baum}%
+        \caption{Baum}\label{fig:cmp_gaussian}%
     \end{center}%
@@
 
-\begin{itemize}
-\item For Lax-Friedrich, $\mu$ affects dissipation
-\item For Local-Lax-Friedrich, $\mu$ affects dissipation much less
-\item Lax-Friedrich approaches Local-Lax-Friedrich as $\mu \to 1.0$
-\item Both methods are unstable for $\mu > 1.0$ due to CFL violation
-\end{itemize}
+A final round of experiments were conducted to compare Lax-Friedrichs to Local Lax Friedrichs.
+These indicate that Local Lax-Friedrichs is decidedly less dissipative and more accurate.
+Figure~\ref{fig:cmp_breakingdam} shows the results of running LLxF with a middling $\mu=0.4$ against LxF with different $\mu$ for the breaking dam scenario.
+
+At $\mu=0.4$, the LLxF method performs only slightly worse than LxF at $\mu=1.0$.
+For any given choice of $\mu$, the LLxF reaches a better solution in the same number of timesteps.
+Unexpectedly, the quality of the solutions converge as $\mu \to 1.0$.
+On the other hand, LLxF requires additional work to calculate $a(x)$ for each cell, though the amount is small and linear in the number of cells.
+
+The stairstep oscillations observed in the breaking dam case for LxF are not present for LLxf.
+This is presumably because the value of $Q_i^n$ enters into the calculation of $Q_i^{n+1}$, albeit rather indirectly, via the max wavespeed $\lambda_{max}$ and then the cell-wise flux $F_{i+1/2}^n$.
 
 \section{Conclusion}
+
+The Lax-Friedrichs method is a simple finite volume technique for solving the shallow water equations.
+Starting from the fundamentals, this paper built up a reference implementation which leverages features of Julia to achieve terseness and clarity not available in languages such as Matlab or C++.
+An abstract finite volume skeleton allows different numerical methods to be swapped in.
+Two variants were implemented, Lax-Friedrichs and Local Lax-Friedrichs.
+The former suffers from excessive numerical viscosity, which the latter largely remedies.
+The causes and consequences of this viscosity were discussed with respect to two model problems, a Gaussian perturbation and a breaking dam.
+It is the author's hope that this document may serve as a concise guide to students exploring this material in the future.
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{IEEEabrv,references}
+\bibliography{IEEEabrv,../references}
 %

### 2017-12-23 paper/paper.tex
Extracted code to appendix which uses minted

@@
 \input{paper-layout}
+\usepackage[outputdir=build]{minted}
+\usepackage{tcolorbox}
+\tcbuselibrary{minted,skins}
+
+\newtcblisting{my_julia}[1][]{%
+  listing engine=minted,
+  minted language=julia,
+  minted options={breaklines,breakanywhere,fontsize=\scriptsize,gobble=8},
+  listing only,
+  before skip=0pt,
+  after skip=8pt,
+  left skip=0pt,
+  right skip=0pt,
+  size=fbox,
+  sharp corners,
+  %colframe=white!75!black,
+  colframe=white,
+  boxrule=0pt,
+  frame hidden,
+  #1
+}
+
 \input{german}
+
 \usepackage{amsmath}
+
 \pagestyle{empty}  % Skip page number on first page
@@
 \begin{itemize}
-\item In the 1-D case, the flow is in a channel of unit width
-\item Horizontal velocity $u(x)$ across any cross section $x$ is constant
-\item Vertical velocity $v(x)$ is small (but not zero) and falls out of the equations
+\item In the 1-D case, the flow is in a channel of unit width.
+\item Horizontal velocity $u(x)$ across any cross section $x$ is constant.
+\item Vertical velocity $v(x)$ is small (but not zero) and falls out of the equations.
 \item Pressure is dominated by hydrostatics: $p=\frac{1}{2}\rho gh^2$
@@
 
-This paper chooses the state variables $q := (h, hu)$ to form a state space, $Q$. This can be expressed in code as follows:
-\begin{verbatim}
-const Float = Float64
-
-mutable struct Q
-    h :: Float
-    hu :: Float
-end
-\end{verbatim}
+This paper chooses the state variables $q := (h, hu)$ to form a state space, $Q$.
+As shown in lines XX-XX, this can be expressed in code as a lightweight mutable structure.
+The choice of single versus double precision for calculations can be specified via a single type alias.
 
@@
 
-This latter problem may be resolved quickly, thanks to Julia's type system.\cite{doi:10.1137/141000671}
+This latter problem may be resolved quickly, thanks to Julia's type system.\cite{doi:10.1137/141000671} Rather than having full objects, Julia provides container-like types.
+Methods have multiple dispatch, which means that a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.
+This means that operations such addition can be extended to custom types in a streamlined way.
+This can be seen in lines XX-XX.
 
-\begin{verbatim}
-import Base.+
-import Base.-
-import Base.*
+Using parametric polymorphism, the spatial domain can be quickly defined as a type alias \verb!QS = Array{Q,1}!
+This states that spatial variables are implemented as a 1-dimensional array with inner type \verb!Q!
+as defined above.
+There is an additional constraint in that there is a ghost cell on either side, so the `true' spatial domain ranges over the indices $[2, n_{cells}+1]$.
 
-function +(lhs::Q, rhs::Q)
-    Q(lhs.h + rhs.h, lhs.hu + rhs.hu)
-end
-
-function -(lhs::Q, rhs::Q)
-    Q(lhs.h - rhs.h, lhs.hu - rhs.hu)
-end
-
-function *(a::Real, q::Q)
-    Q(a*q.h, a*q.hu)
-end
-\end{verbatim}
+Two improvements are possible.
+Firstly, by using \verb!OffsetArray!s instead of regular \verb!Array!s, the indexes can be cheaply translated to $[1, n_{cells}]$.
+Secondly, the type QS can be parameterized by the number of cells, \verb!QS{ncells}!, which would ensure that compatibly sized arrays are always used, that ghost cells always exist, and that the array sizes are known at compile time, allowing for additional compiler optimizations.
+Though potentially useful for larger projects, these features would introduce more complexity than they conceal.
 
@@
 
-This flux function may be expressed in Julia directly:
-
-\begin{verbatim}
-const G = 9.81f0
-function f(q :: Q)
-    Q(q.hu, q.hu^2/q.h + 0.5*G*q.h^2)
-end
-\end{verbatim}
-
-The shallow water equations enter the numerical code in only one other place, the max wave speed function. The physical speeds of wave propagation are given by the eigenvalues of the flux Jacobian,
+This flux function is expressed in lines XX-XX.
+The shallow water equations enter the numerical code in only one other place, the max wave speed function.
+The physical speeds of wave propagation are given by the eigenvalues of the flux Jacobian,
 
@@
 
-will be shown to play an important role in ensuring numerical stability. In Julia it is calculated as:
-
-\begin{verbatim}
-function wavespeed(q :: Q)
-    u = q.hu / q.h
-    c = sqrt(G * q.h)
-    max(abs(u-c), abs(u+c))
-end
-\end{verbatim}
+will be shown to play an important role in ensuring numerical stability. This is implemented on lines XX-XX.
 
@@
 
-Let us now consider a clean organizing principle for the code.
-
-Thanks to parametric polymorphism, the spatial domain can be quickly defined as a type alias \verb!QS = Array{Q,1}!
-This states that spatial variables are implemented as a 1-dimensional array with inner type \verb!Q!
-as defined above.
-There is an additional constraint in that there is a ghost cell on either side, so the `true' spatial domain ranges over the indices $[2, n_{cells}+1]$.
-
-Two improvements are possible.
-Firstly, by using \verb!OffsetArray!s instead of regular \verb!Array!s, the indexes can be cheaply translated to $[1, n_{cells}]$.
-Secondly, the type QS can be parameterized by the number of cells, \verb!QS{ncells}!, which would ensure that compatibly sized arrays are always used, that ghost cells always exist, and that the array sizes are known at compile time, allowing for additional compiler optimizations.
-Though potentially useful for larger projects, these features would introduce more complexity than they conceal.
+Let us now consider a clean organizing principle for the code. A basic skeleton for running experiments is given on lines XX-XX. The following components are allowed to vary:
 
@@
 
-In Julia, the CFL condition is encoded as follows. One useful feature is the function chain operator \verb!|>!, which aids in readability.
-
-\begin{verbatim}
-function cfl_dt(qs::QS, dx::Float, mu::Float)
-    lambda = map(wavespeed,qs) |> maximum
-    dx / lambda * mu
-end
-\end{verbatim}
-
-\section{Model Problems}
+The CFL condition is encoded on lines XX-XX. One useful feature is the function chain operator \verb!|>!, which aids in readability.
 
@@
 \[ F_{i+1/2} = \frac{1}{2} \Big[ f(Q_{i+1}) + f(Q_{i}) - \frac{\Delta x}{\Delta t}(Q_{i+1} - Q_{i})\Big] \]
-
-This can be encoded as a function 
-* Input:  Buffers for $q(x), F_l(x), F_r(x)$, cell and timestep sizes
-* Output: Updated inter-cell flux buffers
-
-\begin{verbatim}
-function lxf!(qs::QS, Fl::QS, Fr::QS, 
-			  ncells::Int, 
-              dx::Float, dt::Float)
-
-    a = dx/dt
-    for x = 2:ncells+2
-        ql,qr = qs[x-1], qs[x]
-        fl,fr = f(ql), f(qr)
-
-        Fr[x-1] = 0.5*((fr-fl) - a*(qr - ql))
-        Fl[x-1] = 0.5*((fr-fl) + a*(qr - ql))
-    end
-end
-\end{verbatim}
 
@@
 The experiment depicted in Figure~\ref{fig:lxf_gaussian_ncells} is rather intuitive.
-A Gaussian perturbation propagates symmetrically across the spatial domain.
+The Gaussian perturbation propagates symmetrically across the spatial domain.
 As the spatial discretization is refined, a shock wave gradually emerges from each (previously symmetric) ripple.
@@
 
-$a_{i-1/2} = \max(\lvert f'(q) \rvert) \ \forall q \in (Q_{i-1}, Q_{i})$
-
-Then $ F_{i-1/2} = \frac{1}{2}[f(Q_{i-1}) + f(Q_i) - a_{i-1/2}(Q_i - Q_{i-1})]$
-
-From the CFL condition, it is clear that this approach is strictly less viscous than basic Lax-Friedrichs:
-
-\[\lvert f'(q)\rvert \leq \frac{\Delta x}{\Delta t}\]
-
 \begin{figure}[h]%
@@
 
-This is represented in code as follows. The flux
+The viscosity parameter $a$ is taken to be the maximum of the wavespeeds of each neighboring cell.
+As shown in Figure~\ref{fig:artificial_viscosity}, this exhibits the desired behavior of being large in regions which are discontinuous or evolving quickly, and being small elsewhere.
 
-\begin{verbatim}
-function llxf!(qs::QS, Fl::QS, Fr::QS, 
-              ncells::Int, 
-              dx::Float, dt::Float)
+\[a_{i-1/2} = \max(\lvert f'(q) \rvert) \ \forall q \in (Q_{i-1}, Q_{i})\]
 
-    for x = 2:ncells+2
-        ql,qr = qs[x-1], qs[x]
-        fl,fr = f(ql), f(qr)
+\[ F_{i-1/2} = \frac{1}{2}[f(Q_{i-1}) + f(Q_i) - a_{i-1/2}(Q_i - Q_{i-1})] \]
 
-        a = max(wavespeed(ql), wavespeed(qr))
+From the CFL condition, it is clear that this approach is strictly less viscous than basic Lax-Friedrichs:
 
-        Fr[x-1] = 0.5*((fr-fl) - a*(qr-ql))
-        Fl[x-1] = 0.5*((fr-fl) + a*(qr-ql))
-    end
-end
-\end{verbatim}
+\[\lvert f'(q)\rvert \leq \frac{\Delta x}{\Delta t}\]
 
@@
 
-%\input{conclusion}
-
 %\section*{Acknowledgment}
@@
 \bibliography{IEEEabrv,../references}
+
+\input{appendix}
+
 %

### 2017-12-23 paper/appendix2.tex
Attractive code listings

\section{Appendix}
\definecolor{bg}{rgb}{0.95,0.95,0.95}

\begin{minted}[linenos,fontsize=\scriptsize,bgcolor=bg]
{julia}

const Float = Float64

"State space datatype"
mutable struct Q
    h :: Float
    hu :: Float
end
\end{minted}

\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=9,bgcolor=bg]
{julia}
# Make Q a vector space

import Base.+, Base.-, Base.*

function +(lhs::Q, rhs::Q)
    Q(lhs.h + rhs.h, lhs.hu + rhs.hu)
end

function -(lhs::Q, rhs::Q)
    Q(lhs.h - rhs.h, lhs.hu - rhs.hu)
end

function *(a::Real, q::Q)
    Q(a*q.h, a*q.hu)
end
\end{minted}

\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=26,bgcolor=bg]
{julia}
const G = 9.81f0

"Shallow-water point-wise flux"
function f(q :: Q)
    Q(q.hu, q.hu^2/q.h + 0.5*G*q.h^2)
end

"Shallow-water point-wise max wave speed"
function wavespeed(q :: Q)
    u = q.hu / q.h
    c = sqrt(G * q.h)
    max(abs(u-c), abs(u+c))
end

\end{minted}

\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=40,bgcolor=bg]
{julia}
"Spatial domain datatype"
QS = Array{Q,1}

"Finite volume skeleton"
function run_experiment(stoptime, ncells::Int, 
    choose_dt, compute_fluxes!, apply_bcs!, apply_ics)

    qs = apply_ics(ncells)
    Fl = Array{Q,1}(ncells+2)
    Fr = Array{Q,1}(ncells+2)

    currenttime = 0
    timesteps = 0
    dx = Float(1000.0/ncells)

    while currenttime < stoptime
        apply_bcs!(qs)
        dt = choose_dt(qs, dx)
        compute_fluxes!(qs, Fl, Fr, ncells, dx, dt)
        update_qs!(qs, Fl, Fr, ncells, dx, dt)
        currenttime += dt
        timesteps += 1
    end
    return (qs, currenttime, timesteps)
end

\end{minted}
\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=67,bgcolor=bg]
{julia}
# Kernels: (qs, Fl, Fr, ncells, dx, dt) -> nothing

"Lax-Friedrichs kernel"
function lxf!(qs::QS, Fl::QS, Fr::QS, 
              ncells::Int, dx::Float, dt::Float)

    a = dx/dt
    for x = 2:ncells+2
        ql,qr = qs[x-1], qs[x]
        fl,fr = f(ql), f(qr)

        Fr[x-1] = 0.5*((fr-fl) - a*(qr - ql))
        Fl[x-1] = 0.5*((fr-fl) + a*(qr - ql))
    end
end

\end{minted}
\newpage
\begin{minted}[breaklines,breakanywhere,linenos,firstnumber=85,fontsize=\scriptsize]
{julia}

"Local Lax-Friedrichs kernel"
function llxf!(qs::QS, Fl::QS, Fr::QS, 
               ncells::Int, dx::Float, dt::Float)

    for x = 2:ncells+2
        ql,qr = qs[x-1], qs[x]
        fl,fr = f(ql), f(qr)

        a = max(wavespeed(ql), wavespeed(qr))

        Fr[x-1] = 0.5*((fr-fl) - a*(qr-ql))
        Fl[x-1] = 0.5*((fr-fl) + a*(qr-ql))
    end
end

# Timestepping strategies: (QS, dx) -> dt

function make_const_dt(dt)
    return (qs, dx) -> dt
end

function cfl_dt(qs::QS, dx::Float, mu::Float)
    lambda = map(wavespeed,qs) |> maximum
    return dx / lambda * mu
end

function make_cfl_dt(correction::Float)
    return (qs, dx) -> cfl_dt(qs, dx, correction)
end

function make_lagging_dt(dt0, correction)
    dt = dt0
    function lagging_dt(qs, dx)
        result = dt
        dt = cfl_dt(qs, dx, correction)
        return result
    end
    return lagging_dt
end

# Initial conditions: ncells -> QS

function gaussian_ics(ncells::Int)
    qs = Array{Q,1}(ncells+2)
    mid = ncells / 2.0 + 1
    for x = 1:ncells+2
        pos_rel = (mid-x) / (ncells/10.0)
        qs[x] = Q(exp(-pos_rel^2) + 1, 0.0)
    end
    return qs
end

function breakingdam_ics(ncells::Int)
    [x < ncells/2+1 ? Q(2,0) : Q(1,0) for x in 1:ncells+2]
end

## Boundary conditions: QS -> nothing

function outflow_bcs!(qs::Array{Q,1})
    qs[1] = qs[2]
    qs[end] = qs[end-1]
end

function periodic_bcs!(qs::Array{Q,1})
    qs[1] = qs[end-1]
    qs[end] = qs[2]
end

function reflecting_bcs!(qs::Array{Q,1})
    qs[1].h = qs[2].h
    qs[1].hu = -qs[2].hu
    qs[end].h = qs[end-1].h
    qs[end].hu = -qs[end-1].hu
end

# Sample experiment run
qs,t,ts = run_experiment(50, 1000, make_cfl_dt(w), 
             lxf!, outflow_bcs!, breakingdam_ics)

\end{minted}

### 2017-12-23 paper/paper.tex
Attractive code listings

@@
 
-\newtcblisting{my_julia}[1][]{%
+% Patch to fix https://github.com/T-F-S/tcolorbox/issues/12
+\makeatletter
+\def\tcb@minted@input@listing#1#2#3#4{%
+  \edef\temp@a{#4}%
+  \ifx\temp@a\@empty%
+  \else%
+    \toks@=\expandafter{#4}%
+    \edef\tcb@temp{\noexpand\usemintedstyle{\the\toks@}}%
+    \tcb@temp%
+  \fi%
+  \toks@=\expandafter{#1}%
+  \edef\tcb@temp{\noexpand\inputminted[\the\toks@]}%
+  \IfFileExists{\minted@outputdir#3}%
+    {\tcb@temp{#2}{\minted@outputdir#3}}%
+    {\tcb@temp{#2}{#3}}%
+}
+\makeatother
+
+\newtcblisting{myjulia}{%
   listing engine=minted,
   minted language=julia,
-  minted options={breaklines,breakanywhere,fontsize=\scriptsize,gobble=8},
+  minted options={fontsize=\scriptsize,firstnumber=last,linenos},
   listing only,
@@
   size=fbox,
-  sharp corners,
-  %colframe=white!75!black,
-  colframe=white,
+  rounded corners,
+  colframe=white!75!black,
   boxrule=0pt,
   frame hidden,
-  #1
+  width=\linewidth - 1em
 }
@@
 \bibliography{IEEEabrv,../references}
-
-\input{appendix}
+\newpage
+\input{appendix3}
 

### 2017-12-29 paper/paper.tex
Another round of changes to the paper

@@
 
-\pagestyle{empty}  % Skip page number on first page
+\pagestyle{plain}  % Skip page number on first page
 
@@
 
-%\input{abstract}
+\begin{abstract}
+Lax-Friedrichs and Local Lax-Friedrichs are two numerical methods for solving the shallow water equations. 
+1 sentence intro
+2 sentences methods
+1 sentence impl/results
+why do people care?
+\end{abstract}
 
-%\begin{keywords}
-%	Shallow water equations, Finite volumes, Lax-Friedrichs method, Rusanov, artificial viscosity, Julia
-%\end{keywords}
+\begin{keywords}
+Shallow water equations, Finite volumes, Lax-Friedrichs method, Rusanov, artificial viscosity, Julia
+\end{keywords}
 
@@
 
-This paper has three aims.
-Firstly, it shall introduce the shallow water equations and describe how finite volume methods can be used to compute a numerical solution.
-Secondly, it shall explore two such methods, the Lax-Friedrich and the Local-Lax-Friedrich (Rusanov) methods, describing them conceptually via equations and concretely via code snippets.
-The code is organized to map to the concepts as cleanly as possible and to leverage higher-order programming techniques and patterns.
-Finally, the code shall be used to generate solutions to several model problems, which can be examined in order to gain insight into the concept and consequences of numerical viscosity.
+This paper introduces two finite volume methods for solving the shallow water equations.
+These methods are applied to solving two model problem: the Gaussian perturbation and the breaking dam.
+The simplest convergent method, Lax-Friedrichs, captures the desired dynamics but exhibits nonphysical viscous behavior.
+An analysis shows that this artificial viscosity arises from the method itself.
+A second method, Local Lax-Friedrichs (Rusanov), is shown via analysis and experiment to improve on the former method.
+A concise Julia implementation is provided and described.
 
@@
 
-This paper chooses the state variables $q := (h, hu)$ to form a state space, $Q$.
-As shown in lines XX-XX, this can be expressed in code as a lightweight mutable structure.
-The choice of single versus double precision for calculations can be specified via a single type alias.
-
-There are several tradeoffs associated with this representation of state space.
-In performance-critical code, the choice of data structure is driven primarily by memory access patterns, which are not considered here.
-The principal benefit of this representation is that it moves some of the semantics of the problem domain into the type system, which would permit future extensions such as awareness of physical units.
-On the other hand, this representation is unaware of the fact that $Q$ forms a vector space, resulting in duplicated, messier, and possibly less efficient code.
-
-This latter problem may be resolved quickly, thanks to Julia's type system.\cite{doi:10.1137/141000671} Rather than having full objects, Julia provides container-like types.
-Methods have multiple dispatch, which means that a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.
-This means that operations such addition can be extended to custom types in a streamlined way.
-This can be seen in lines XX-XX.
-
-Using parametric polymorphism, the spatial domain can be quickly defined as a type alias \verb!QS = Array{Q,1}!
-This states that spatial variables are implemented as a 1-dimensional array with inner type \verb!Q!
-as defined above.
-There is an additional constraint in that there is a ghost cell on either side, so the `true' spatial domain ranges over the indices $[2, n_{cells}+1]$.
-
-Two improvements are possible.
-Firstly, by using \verb!OffsetArray!s instead of regular \verb!Array!s, the indexes can be cheaply translated to $[1, n_{cells}]$.
-Secondly, the type QS can be parameterized by the number of cells, \verb!QS{ncells}!, which would ensure that compatibly sized arrays are always used, that ghost cells always exist, and that the array sizes are known at compile time, allowing for additional compiler optimizations.
-Though potentially useful for larger projects, these features would introduce more complexity than they conceal.
-
-The shallow water equations may be defined over this state space in an abstract way. Because they match the pattern of a general conservation law,
+The shallow water equations may be defined for a state vector $q := (h,hu)$ in an abstract way, permitting a decoupling between the underlying equation and the numerical method.
+Because the shallow-water equtions match the pattern of a general conservation law,
 
@@
 
-their dynamics may be captured as a flux function, $f$, such that
+the dynamics are captured by the flux function $f$, where
 
@@
 
-This flux function is expressed in lines XX-XX.
+This flux function is coded in lines 27--30.
 The shallow water equations enter the numerical code in only one other place, the max wave speed function.
-The physical speeds of wave propagation are given by the eigenvalues of the flux Jacobian,
+The system's wavespeeds are given by the eigenvalues of the flux Jacobian,
 
@@
 
-will be shown to play an important role in ensuring numerical stability. This is implemented on lines XX-XX.
+will be shown to play an important role in ensuring numerical stability. The corresponding implementation is provided on lines 32--37.
 
@@
 
-Finite volumes are a family of numerical methods which are well suited for computing an approximate solution to hyperbolic systems such as the shallow water equations.\cite{leveque2002finite} They have two main advantages over other methods, such as finite differences and finite elements.
+Finite volumes are a family of numerical methods which are well suited for computing an approximate solution to hyperbolic systems such as the shallow water equations.\cite{leveque2002finite} They have two main advantages over methods such as finite elements.
 Firstly, they automatically conserve physical quantities $q$.
 Secondly, they can capture discontinuities, such as shocks, which is considerably simpler than explicitly tracking them.
-The main downside is that the profile of the shock will be smeared by the spatial discretization.
+One downside is that the profile of the shock will be smeared by the spatial discretization.
 
@@
 
-Since these equations hold for an arbitrary control volume, it makes sense to partition the spatial domain into cells.
+Since these equations hold for an arbitrary control volume, we partition the spatial domain into cells, rather than points.
 For simplicity these cells are assumed to be regular, although the equations hold just as well over an unstructured mesh.
@@
 
-\begin{multline}
-\int_{C_i} q(x,t_{n+1})dx - \int_{C_i} q(x,t_n) dx = \\ \int_{t_n}^{t_{n+1}} f(q(x_{i-1/2},t))dt - \int_{t_n}^{t_{n+1}} f(q(x_{i+1/2},t))dt
-\end{multline}
+\begin{align*}
+\int_{C_i} q(x,t_{n+1})\,dx & - \int_{C_i} q(x,t_n)\,dx = \\ \int_{t_n}^{t_{n+1}} f(q(x_{i-1/2},t))\,dt & - \int_{t_n}^{t_{n+1}} f(q(x_{i+1/2},t))\,dt
+\end{align*}
 
 \begin{itemize}
-\item Let $Q_{i}^{n} \approx \frac{1}{\Delta x} \int_{C_i} q(x,t_n) dx $ be the approximate average $q$ in cell $i$ at time $n$.
+\item Let $Q_{i}^{n} \approx \frac{1}{\Delta x} \int_{C_i} q(x,t_n)\,dx $ be the approximate average $q$ in cell $i$ at time $n$.
 
-\item Let $F_{i+1/2}^{n} \approx \frac{1}{\Delta t} \int_{t_n}^{t_{n+1}} f(q(x_{i+1/2},t))dt$ be the approximate average flux $f$ from cell $i$ into $i+1$.
+\item Let $F_{i+1/2}^{n} \approx \frac{1}{\Delta t} \int_{t_n}^{t_{n+1}} f(q(x_{i+1/2},t))\,dt$ be the approximate average flux $f$ from cell $i$ into $i+1$.
 \end{itemize}
@@
 
-Ideally, the flux function can be calculated using only the state in the adjacent cells, which yields a fully discrete method:
+Ideally, the cell-wise flux function can be calculated using only the state in the adjacent cells, which yields a fully discrete method:
 
@@
 Similarly, they hold for any choice of hyperbolic PDE, which is abstracted away as a state space $Q_i$, a pointwise flux $f(Q_i)$ and max wavespeed $\lambda_{max}(Q_i)$.
-Thus, the implementation may be organized into an abstract finite volume method which is parameterized with different strategies.
-
-\section{Code architecture}
-
-Let us now consider a clean organizing principle for the code. A basic skeleton for running experiments is given on lines XX-XX. The following components are allowed to vary:
-
-\begin{itemize}
-    \item Flux function f
-    \item Initial conditions
-    \item Boundary conditions
-    \item Timestep size
-\end{itemize}
-
-In traditional object-oriented programming, this decoupling is usually achieved via the Strategy Pattern. Because Julia has first-class functions (lambdas), it can be achieved much more tersely.
+This natural decoupling of the mathematical machinery can be directly exploited in the implementation by parameterizing a generic finite volume skeleton with different strategies.
 
@@
 Because this code uses the same timestep size for all cells, the maximum wave speed must be calculated across the entire spatial domain for each timestep iteration.
-The maximum wavespeed calculated from the flux Jacobian is exact only in the small-perturbation case, so an additional correction factor, $\mu$, scales the timestep down.
+The maximum wavespeed calculated from the flux Jacobian is exact only in the small-perturbation case, so an additional correction factor, $\mu \in [0,1]$, scales the timestep down.
 Thus the experimenter has two knobs for balancing speed, accuracy, and stability: The spatial discretization $\Delta x$ and the timestep correction $\mu$.
 
-The CFL condition is encoded on lines XX-XX. One useful feature is the function chain operator \verb!|>!, which aids in readability.
+The CFL condition is encoded on lines 93-96.
 
@@
 
-Having built up a skeleton of a finite volume method, it is finally time to furnish the most essential missing piece, the cell-wise flux function.
-The most obvious choice of $F$ is simply the average of the fluxes at the center of either cell.
+Different finite-volume methods may be derived from a common skeleton by providing a choice of $\mathcal F$.
+The simplest choice is simply the average of the fluxes at the center of either neighboring cell.
 
-\[ F_{i+1/2} = \frac{1}{2}\big[f(Q_{i} + f(Q_{i+1})\big] \]
+\[ F_{i+1/2} = \frac{1}{2}\big[f(Q_{i}) + f(Q_{i+1})\big] \]
 
@@
 
-The resulting update scheme becomes:
+The resulting update scheme becomes
 
@@
 
-which may be rearranged to yield a cell-wise flux function:
+This may be rearranged to yield a cell-wise flux function:
 
@@
 
-The Lax-Friedrichs method implemented was run against two model problems.
+The Lax-Friedrichs flux function is implemented on lines 60--72.
+It was run against two model problems.
 The first used a Gaussian perturbation for its initial condition with reflecting boundary conditions.
@@
 It should be noted that these shock waves are smeared across tens or hundreds of cells, far more than the finite volume method presumably requires.
-Furthermore, the crest is reduced and the troughs are filled in: Decreasing the spatial discretization appears to increase dissipation.
+Furthermore, the crest is reduced and the troughs are filled in: coarsening the mesh increases dissipation.
 
-Applying the same experiment to the breaking dam scenario, shown in Figure~\ref{fig:lxf_breakingdam_ncells} yields similar observations, with an unexpected twist.
+Applying the same experiment to the breaking dam scenario, as shown in Figure~\ref{fig:lxf_breakingdam_ncells}, yields an additional, unexpected twist.
 As the mesh becomes coarser, a nonphysical stairstep pattern emerges.
@@
 The key idea behind artificial viscosity is that an approximate solution to a certain problem is oftentimes an exact solution to an approximation of that problem.
-The trick is to regard simple discretizatons of a PDE as being building blocks which can be assembled into more sophisticated (and correspondingly more approximate) discretizations.
+The trick is to regard simple discretizatons of PDEs as being building blocks which can be assembled into more sophisticated (and correspondingly more approximate) discretizations.
 One can then map the sophisticated discretization directly back to a modified PDE.
@@
 
-Local-Lax-Friedrichs is a modification which consciously uses artificial viscosity in a clever way.
+Local-Lax-Friedrichs is a modification which retains the stability benefits of artificial viscosity while decreasing the resulting dissipation.
 In the basic Lax-Friedrichs, the viscosity parameter $a$ is a constant derived from the timestep $\Delta t$, which in turn is derived from the CFL condition.
 Although the timestep is contrained to be constant over the entire spatial domain (local time stepping is not considered in the scope of this paper), the viscosity is not.
-On the contrary, varying the viscosity over the spatial domain is advantageous: in regions with shocks and other discontinuities, the presence of viscosity helps ensure numerical stability, and in smooth regions, the lack of viscosity leads to a less dissipative and hence more accurate solution.
+Varying the viscosity over the spatial domain is advantageous: in regions with shocks or other discontinuities, the presence of viscosity improves numerical stability, and in smooth regions, the lack of viscosity leads to a less dissipative and hence more accurate solution.
 
@@
 
-The viscosity parameter $a$ is taken to be the maximum of the wavespeeds of each neighboring cell.
+The viscosity parameter $a$ is taken to be the maximum of the wavespeeds of either neighboring cell.
 As shown in Figure~\ref{fig:artificial_viscosity}, this exhibits the desired behavior of being large in regions which are discontinuous or evolving quickly, and being small elsewhere.
@@
 The Local Lax-Friedrichs method does not entirely break the coupling between dissipation and $\mu$, but it strongly reduces it, as depicted in Figure~\ref{fig:llxf_breakingdam_mu}.
-This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, only for LLxF.
+This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, substituting only the LxF kernel (as implemented in lines 60--72) with the LLxF kernel (as implemented in lines 73--86).
 \begin{figure}[h]%
@@
         \includegraphics[scale=0.6]{../plots/llxf_breakingdam_mu.png}%
-        \caption{Baum}\label{fig:llxf_breakingdam_mu}%
+        \caption{Local Lax-Friedrichs, breaking dam scenario. }\label{fig:llxf_breakingdam_mu}%
     \end{center}%
@@
         \includegraphics[scale=0.6]{../plots/lxf_llxf_breakingdam.png}%
-        \caption{Baum}\label{fig:cmp_breakingdam}%
+        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, breakin dam scenario}\label{fig:cmp_breakingdam}%
     \end{center}%
@@
         \includegraphics[scale=0.6]{../plots/lxf_llxf_gaussian.png}%
-        \caption{Baum}\label{fig:cmp_gaussian}%
+        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, gaussian perturbation scenario.}\label{fig:cmp_gaussian}%
     \end{center}%
@@
 For any given choice of $\mu$, the LLxF reaches a better solution in the same number of timesteps.
-Unexpectedly, the quality of the solutions converge as $\mu \to 1.0$.
+The quality of the solutions converge as $\mu \to 1.0$.
 On the other hand, LLxF requires additional work to calculate $a(x)$ for each cell, though the amount is small and linear in the number of cells.
@@
 This is presumably because the value of $Q_i^n$ enters into the calculation of $Q_i^{n+1}$, albeit rather indirectly, via the max wavespeed $\lambda_{max}$ and then the cell-wise flux $F_{i+1/2}^n$.
+
+\section{Implementation}
+
+This final section describes the design decisions behind the sample code provided in the appendix.
+The code is broken into chunks corresponding to the theoretical development in the preceding sections.
+It is written in Julia, a language designed for numerical programming.
+The choice of Julia enabled a particularly terse code representation which still maps closely to the theory.
+
+The first design decision is the choice of data representation.
+Julia eschews standard object-oriented idioms in favor of a more functional and data-driven approach.
+The basic composite type is a simple structure, immutable by default, with restricted inheritance and encapsulation.
+
+The state space $Q$ is a mutable structure.
+The spatial domain $q(x) \forall x$ is defined as a one-dimensional array with inner type $Q$.
+The spatial domain has a ghost cell on either side to enforce boundary conditions, so the physical cell indices range over $[2, n_{cells}+1]$.
+
+These additional constraints can be expressed within the type system.
+Firstly, by using \verb!OffsetArray!s instead of regular \verb!Array!s, the indices can be cheaply translated to $[1, n_{cells}]$.
+Secondly, one may parameterize the spatial domain by the number of cells, \verb|QS{ncells}|, e.g.
+by using \verb|StaticArray|s, ensuring that the array size is known at JIT-compile time.
+This allows for array-size correctness checks and compiler optimizations.
+Finally, a custom constructor could encapsulate all of this and enforce the existence of the ghost cells.
+These features may prove helpful on larger projects, though they introduce additional complexity.
+
+There are several tradeoffs associated with using an array-of-structures compared to a structure-of-arrays.
+In performance-critical code, this choice is driven primarily by memory access patterns, which are not considered here.
+The principal benefit of this representation is that it moves more of the semantics of the problem domain into the type system.
+On the other hand, this representation is initially unaware of the fact that $Q$ forms a vector space, resulting in duplicated, messier, and possibly less efficient code.
+
+This problem may be resolved quickly.
+Methods in Julia have multiple dispatch, which means that a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.\cite{doi:10.1137/141000671} This means that operations such addition can be extended to custom types in a unified way.
+Lines 12--26 extend addition and scalar multiplication over the state space type $Q$.
+
+Let us now consider a clean organizing principle for the code.
+A basic skeleton for running experiments is given on lines 38--59.
+The following components are allowed to vary: Flux function f, initial conditions, boundary conditions, timestep size.
+In traditional object-oriented programming, this decoupling is usually achieved via the Strategy Pattern.
+Because Julia has first-class functions (lambdas), it can be achieved much more tersely.
 

### 2018-01-13 paper/paper.tex
Straggler changes to the paper draft

@@
 Lax-Friedrichs and Local Lax-Friedrichs are two numerical methods for solving the shallow water equations.
-1 sentence intro
-2 sentences methods
-1 sentence impl/results
-why do people care?
+This paper introduces both and provides a concise implementation in Julia.
+The Gaussian perturbation and breaking dam model problems are solved for each method over a range of parameters.
+The cause and effects of artificial viscosity are explored analytically and experimentally.
 \end{abstract}
@@
 
-The CFL condition is encoded on lines 93-96.
+The CFL condition is encoded on lines 95--98.
 
@@
 
-The Lax-Friedrichs flux function is implemented on lines 60--72.
+The Lax-Friedrichs flux function is implemented on lines 62--74.
 It was run against two model problems.
@@
 
-\begin{figure}[h]%
+\begin{figure}[t]%
     \begin{center}%
@@
         \caption{Wavespeed over the spatial domain}\label{fig:artificial_viscosity}%
+    \end{center}%
+\end{figure}
+
+\begin{figure}[t]%
+    \begin{center}%
+        \includegraphics[scale=0.6]{../plots/llxf_breakingdam_mu.png}%
+        \caption{Local Lax-Friedrichs, breaking dam scenario. }\label{fig:llxf_breakingdam_mu}%
     \end{center}%
@@
 The Local Lax-Friedrichs method does not entirely break the coupling between dissipation and $\mu$, but it strongly reduces it, as depicted in Figure~\ref{fig:llxf_breakingdam_mu}.
-This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, substituting only the LxF kernel (as implemented in lines 60--72) with the LLxF kernel (as implemented in lines 73--86).
-\begin{figure}[h]%
-    \begin{center}%
-        \includegraphics[scale=0.6]{../plots/llxf_breakingdam_mu.png}%
-        \caption{Local Lax-Friedrichs, breaking dam scenario. }\label{fig:llxf_breakingdam_mu}%
-    \end{center}%
-\end{figure}
+This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, substituting only the LxF kernel (as implemented in lines 62--74) with the LLxF kernel (as implemented in lines 75--88).
 
 \section{LxF vs LLxF Comparison}
-
-\begin{figure}[h]%
-    \begin{center}%
-        \includegraphics[scale=0.6]{../plots/lxf_llxf_breakingdam.png}%
-        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, breakin dam scenario}\label{fig:cmp_breakingdam}%
-    \end{center}%
-\end{figure}
-\begin{figure}[h]%
-    \begin{center}%
-        \includegraphics[scale=0.6]{../plots/lxf_llxf_gaussian.png}%
-        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, gaussian perturbation scenario.}\label{fig:cmp_gaussian}%
-    \end{center}%
-\end{figure}
 
@@
 This is presumably because the value of $Q_i^n$ enters into the calculation of $Q_i^{n+1}$, albeit rather indirectly, via the max wavespeed $\lambda_{max}$ and then the cell-wise flux $F_{i+1/2}^n$.
+
+\begin{figure}[t]%
+    \begin{center}%
+        \includegraphics[scale=0.6]{../plots/lxf_llxf_breakingdam.png}%
+        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, breakin dam scenario}\label{fig:cmp_breakingdam}%
+    \end{center}%
+\end{figure}
+
+\begin{figure}[t]%
+    \begin{center}%
+        \includegraphics[scale=0.6]{../plots/lxf_llxf_gaussian.png}%
+        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, gaussian perturbation scenario.}\label{fig:cmp_gaussian}%
+    \end{center}%
+\end{figure}
 
@@
 
-There are several tradeoffs associated with using an array-of-structures compared to a structure-of-arrays.
-In performance-critical code, this choice is driven primarily by memory access patterns, which are not considered here.
-The principal benefit of this representation is that it moves more of the semantics of the problem domain into the type system.
-On the other hand, this representation is initially unaware of the fact that $Q$ forms a vector space, resulting in duplicated, messier, and possibly less efficient code.
-
-This problem may be resolved quickly.
-Methods in Julia have multiple dispatch, which means that a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.\cite{doi:10.1137/141000671} This means that operations such addition can be extended to custom types in a unified way.
+The choice of an array-of-structures representation is most consistent with the theory, but disregards the useful fact that $Q$ forms a vector space.
+This problem is readily resolved thanks to \emph{multiple dispatch}.
+In Julia, a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.\cite{doi:10.1137/141000671} This means that operations such addition can be extended to custom types in a unified way.
 Lines 12--26 extend addition and scalar multiplication over the state space type $Q$.
 
-Let us now consider a clean organizing principle for the code.
-A basic skeleton for running experiments is given on lines 38--59.
-The following components are allowed to vary: Flux function f, initial conditions, boundary conditions, timestep size.
-In traditional object-oriented programming, this decoupling is usually achieved via the Strategy Pattern.
-Because Julia has first-class functions (lambdas), it can be achieved much more tersely.
+The basic finite volume skeleton is given on lines 38--61.
+The skeleton accepts abstract strategies for flux computations, timestep computations, initial conditions, and boundary conditions.
+These strategies are implemented as first-class functions rather than as an object heirarchy.
+Lines 91--93 demonstrate a higher-order function which acts like a constructor for timestep strategies.
+Lines 100--102 demonstrate \emph{currying} to convert a three-argument \verb!cfl_dt()!
+into a two-argument version.
+Lines 104--112 demonstrate using the method's \emph{closure} to store arbitrary state, in this case the $\Delta t$ calculated during the preceding function call.
+Finally, lines 147--148 ties everything together to run Lax-Friedrichs on a breaking dam scenario.
 

### 2018-01-13 llxf.jl
Separated Julia code from Jupyter notebook, verified that it runs

@@ -1,0 +1,159 @@
+# Define precision
+const Float = Float64
+
+# Define state space datatype
+mutable struct Q
+    h :: Float
+    hu :: Float
+end
+
+# Define spatial domain datatype
+QS = Array{Q,1}
+
+# Make Q a vector space
+
+import Base.+, Base.-, Base.*
+
+function +(lhs::Q, rhs::Q)
+    Q(lhs.h + rhs.h, lhs.hu + rhs.hu)
+end
+
+function -(lhs::Q, rhs::Q)
+    Q(lhs.h - rhs.h, lhs.hu - rhs.hu)
+end
+
+function *(a::Real, q::Q)
+    Q(a*q.h, a*q.hu)
+end
+
+"Shallow-water point-wise flux"
+const G = 9.81f0
+function f(q :: Q)
+    Q(q.hu, q.hu^2/q.h + 0.5*G*q.h^2)
+end
+
+"Shallow-water point-wise max wave speed"
+function wavespeed(q :: Q)
+    u = q.hu / q.h
+    c = sqrt(G * q.h)
+    max(abs(u-c), abs(u+c))
+end
+
+"Finite volume skeleton"
+function run(stoptime::Real, ncells::Int, choose_dt,
+             compute_fluxes!, apply_bcs!, apply_ics)
+
+    qs = apply_ics(ncells)
+    Fl = QS(ncells+2)
+    Fr = QS(ncells+2)
+
+    currenttime = 0
+    timesteps = 0
+    dx = Float(1000.0/ncells)
+    
+    while currenttime < stoptime
+        apply_bcs!(qs)
+        dt = choose_dt(qs, dx)
+        compute_fluxes!(qs, Fl, Fr, ncells, dx, dt)
+        for x = 2:ncells+1
+            qs[x] -= dt/dx * (Fr[x] + Fl[x-1])
+        end
+        currenttime += dt
+        timesteps += 1
+    end
+    return (qs, currenttime, timesteps)
+end
+
+"Lax-Friedrichs kernel"
+function lxf!(qs::QS, Fl::QS, Fr::QS, 
+              ncells::Int, dx::Float, dt::Float)
+    
+    a = dx/dt
+    for x = 2:ncells+2
+        ql,qr = qs[x-1], qs[x]
+        fl,fr = f(ql), f(qr)        
+
+        Fr[x-1] = 0.5*((fr-fl) - a*(qr - ql))
[…]

### 2018-01-13 src/WavePropagation.cpp
Removed obsolete files

@@ -1,144 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer <rannabau@in.tum.de>
- */
-
-#include "WavePropagation.h"
-#include "iostream"
-#include <math.h>
[…]

### 2018-01-13 src/WavePropagation.h

@@ -1,148 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Ranbauer <rannabau@in.tum.de>
- */
-
-#ifndef WAVEPROPAGATION_H_
-#define WAVEPROPAGATION_H_
-
-#include "types.h"
-
-// Include the solver that is currently used
-#define SUPPRESS_SOLVER_DEBUG_OUTPUT
-
-/**
- * Allocated variables:
- *   unknown h are defined on grid indices [0,..,n+1] (done by the caller)
- *     -> computational domain is [1,..,nx]
[…]

### 2018-01-13 src/main.cpp

@@ -1,107 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer <rannabau@in.tum.de>
- */
-
-#include "types.h"
-#include "WavePropagation.h"
-#include "scenarios/gaussian.h"
-#include "scenarios/breakingdam.h"
-#include "writer/VtkWriter.h"
-#include "tools/args.h"
-
-#include <cstring>
-
-int main(int argc, char** argv)
-{
-	// Parse command line parameters
-	tools::Args args(argc, argv);
-
-	// Scenario
[…]

### 2018-01-13 src/scenarios/breakingdam.h

@@ -1,86 +1,0 @@
-/**
- * @file
- *  This file is part of SWE1D
- *
- *  SWE1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  SWE1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with SWE1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von SWE1D.
- *
- *  SWE1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  SWE1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2013 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- */
-
-#ifndef SCENARIOS_BREAKINGDAM_H_
-#define SCENARIOS_BREAKINGDAM_H_
-
-#include "types.h"
-#include <iostream>
-#include <math.h>
-
-namespace scenarios
-{
-
-class BreakingDam
-{
-private:
-	/** Number of cells */
-	const unsigned int m_size;
-
-public:
- BreakingDam(unsigned int size)
[…]

### 2018-01-13 src/scenarios/gaussian.h

@@ -1,81 +1,0 @@
-/**
- * @file
- *  This file is part of SWE1D
- *
- *  SWE1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  SWE1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with SWE1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von SWE1D.
- *
- *  SWE1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  SWE1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2013 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- */
-
-#ifndef SCENARIOS_GAUSSIAN_H_
-#define SCENARIOS_GAUSSIAN_H_
-
-#include "types.h"
-#include <iostream>
-#include <math.h>
-
-namespace scenarios
-{
-
-class Gaussian
-{
-private:
-	/** Number of cells */
-	const unsigned int m_size;
-
-public:
-	Gaussian(unsigned int size)
[…]

### 2018-01-13 src/tools/args.cpp

@@ -1,40 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- *
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer    <rannabau@in.tum.de>
- */
-
-#include "args.h"
-
-

### 2018-01-13 src/tools/args.h

@@ -1,136 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- *
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer    <rannabau@in.tum.de>
- */
-
-#ifndef TOOLS_ARGS_H_
-#define TOOLS_ARGS_H_
-
-#include "tools/logger.h"
-
-#include <getopt.h>
-#include <cstdlib>
-#include <iostream>
-#include <sstream>
-
-namespace tools
-{
-
-/**
- * Parse command line arguments
- */
-class Args
-{
-private:
-	/** Domain size */
-	unsigned int m_size;
[…]

### 2018-01-13 src/tools/logger.cpp

@@ -1,40 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer    <rannabau@in.tum.de>
- */
-
-#include "logger.h"
-
-tools::Logger tools::Logger::logger;

### 2018-01-13 src/tools/logger.h

@@ -1,155 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- *
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer    <rannabau@in.tum.de>
- */
-
-
-#ifndef TOOLS_LOGGER_H_
-#define TOOLS_LOGGER_H_
-
-#include <cstdlib>
-#include <iostream>
-
-namespace tools {
-
-class Logger {
-public:
-	enum Level { INFO, WARNING, ERROR };
-
-private:
-	/** Stream where we print all message */
-	std::ostream* m_output;
-
-private:
-	Logger()
-		: m_output(&std::cout)
-	{
-	}
-
-public:
[…]

### 2018-01-13 src/types.h

@@ -1,53 +1,0 @@
-/**
- * @file
- *  This file is part of SWE1D
- *
- *  SWE1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  SWE1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with SWE1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von SWE1D.
- *
- *  SWE1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  SWE1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- *
- * @copyright 2013 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- */
-
-#ifndef TYPES_H_
-#define TYPES_H_
-
-// Switch precision here
-typedef float T;
-//typedef double T;
-
-struct Q {
-	T h;
-	T hu;
-
-  //	Q operator-(Q & lhs, Q & rhs) {
-	//	return {lhs.h - rhs.h, lhs.hu - rhs.hu};
-	//}
-};
-
-#endif /* TYPES_H_ */

### 2018-01-13 src/writer/ConsoleWriter.h

@@ -1,82 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer    <rannabau@in.tum.de>
- */
-
-#ifndef CONSOLEWRITER_H_
-#define CONSOLEWRITER_H_
-
-#include "types.h"
-
-#include <iostream>
-
-namespace writer
-{
-
-/**
- * A simple writer class, that writes h and hu to stdout (or another ostream)
- */
-class ConsoleWriter
-{
-private:
-	std::ostream &m_ostream;
-
-public:
[…]

### 2018-01-13 src/writer/VtkWriter.h

@@ -1,193 +1,0 @@
-/**
- * @file
- *  This file is part of ADV1D
- *
- *  ADV1D is free software: you can redistribute it and/or modify
- *  it under the terms of the GNU General Public License as published by
- *  the Free Software Foundation, either version 3 of the License, or
- *  (at your option) any later version.
- *
- *  ADV1D is distributed in the hope that it will be useful,
- *  but WITHOUT ANY WARRANTY; without even the implied warranty of
- *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *  GNU General Public License for more details.
- *
- *  You should have received a copy of the GNU General Public License
- *  along with ADV1D.  If not, see <http://www.gnu.org/licenses/>.
- *
- *  Diese Datei ist Teil von ADV1D.
- *
- *  ADV1D ist Freie Software: Sie koennen es unter den Bedingungen
- *  der GNU General Public License, wie von der Free Software Foundation,
- *  Version 3 der Lizenz oder (nach Ihrer Option) jeder spaeteren
- *  veroeffentlichten Version, weiterverbreiten und/oder modifizieren.
- *
- *  ADV1D wird in der Hoffnung, dass es nuetzlich sein wird, aber
- *  OHNE JEDE GEWAEHELEISTUNG, bereitgestellt; sogar ohne die implizite
- *  Gewaehrleistung der MARKTFAEHIGKEIT oder EIGNUNG FUER EINEN BESTIMMTEN
- *  ZWECK. Siehe die GNU General Public License fuer weitere Details.
- *
- *  Sie sollten eine Kopie der GNU General Public License zusammen mit diesem
- *  Programm erhalten haben. Wenn nicht, siehe <http://www.gnu.org/licenses/>.
- * 
- * @copyright 2017 Technische Universitaet Muenchen
- * @author Sebastian Rettenberger <rettenbs@in.tum.de>
- * @author Leonhard Rannabauer    <rannabau@in.tum.de>
- */
-
-#ifndef VTKWRITER_H_
-#define VTKWRITER_H_
-
-#include "types.h"
-
-#include <cassert>
-#include <fstream>
-#include <sstream>
-#include <string>
-
-namespace writer
-{
-
-/**
- * A writer class that generates vtk files
- */
-class VtkWriter
-{
-private:
[…]

### 2018-01-15 paper/paper.tex
Numbered equations, removed unnecessary refs to code, made language use less ponderous and awkward, freed up some space

@@
 
-This paper introduces two finite volume methods for solving the shallow water equations.
-These methods are applied to solving two model problem: the Gaussian perturbation and the breaking dam.
-The simplest convergent method, Lax-Friedrichs, captures the desired dynamics but exhibits nonphysical viscous behavior.
-An analysis shows that this artificial viscosity arises from the method itself.
-A second method, Local Lax-Friedrichs (Rusanov), is shown via analysis and experiment to improve on the former method.
-A concise Julia implementation is provided and described.
+This paper introduces two finite volume methods for solving the shallow water equations, and applies them to the Gaussian perturbation and breaking dam problems.
+The first three sections respectively introduce the shallow water equations, the general finite volume method, and the CFL condition.
+Section 5 derives the Lax-Friedrichs method, and Section 6 demonstrates that it captures the desired dynamics but exhibits undesired viscous behavior.
+Section 7 shows that this viscosity arises from the method's numerics.
+Section 8 introduces the Local Lax-Friedrichs (Rusanov) method, and Section 9 shows that the viscous behavior is largely mitigated.
+Section 10 discusses a concise Julia implementation, which is provided in the Appendix.
 
@@
 The shallow water equations may be defined for a state vector $q := (h,hu)$ in an abstract way, permitting a decoupling between the underlying equation and the numerical method.
-Because the shallow-water equtions match the pattern of a general conservation law,
+Because the shallow-water equtions match the pattern of a general conservation law, $q_t(x,t) + f(q(x,t))_x = 0$, the dynamics are captured by the pointwise flux function $f$, where
 
-\[q_t(x,t) + f(q(x,t))_x = 0\]
+\begin{equation}
+\label{eq:flux}
+f : (q_1, q_2) \mapsto \Big(q_2, \frac{q_{2}^{2}}{q_{1}} + \frac{gq_1^2}{2}\Big)
+\end{equation}
 
-the dynamics are captured by the flux function $f$, where
-
-\[f : (q_1, q_2) \mapsto \begin{bmatrix} q_2\\ q_2^2/q_1 + \frac{1}{2}gq_1^2\end{bmatrix}\]
-
-This flux function is coded in lines 27--30.
-The shallow water equations enter the numerical code in only one other place, the max wave speed function.
-The system's wavespeeds are given by the eigenvalues of the flux Jacobian,
+The shallow water equations enter the method in only one other place, the max wave speed function. The system's wavespeeds are given by the eigenvalues of the flux Jacobian,
 
@@
 
-\[\lambda_{max} = \max\ \lvert u \pm \sqrt{gh} \rvert \]
+\begin{equation}
+\label{eq:wavespeed}
+\lambda_{max} = \max\ \lvert u \pm \sqrt{gh} \rvert
+\end{equation}
 
-will be shown to play an important role in ensuring numerical stability. The corresponding implementation is provided on lines 32--37.
+will be shown to play an important role in ensuring numerical stability.
 
@@
 
-\[Q_{i}^{n+1} = Q_{i}^{n} - \frac{\Delta t}{\Delta x}(F_{i+1/2}^{n} - F_{i-1/2}^{n}) \]
+\begin{equation}
+\label{eq:fv}
+Q_{i}^{n+1} = Q_{i}^{n} - \frac{\Delta t}{\Delta x}(F_{i+1/2}^{n} - F_{i-1/2}^{n})
+\end{equation}
 
-Ideally, the cell-wise flux function can be calculated using only the state in the adjacent cells, which yields a fully discrete method:
+When the cellwise flux function can be calculated using only the state in the adjacent cells, the method is fully discrete:
 
@@
 
-Taking a step back, it is clear that these also fall under the umbrella of finite difference methods.
+Taking a step back, it is clear that these methods also fall under the umbrella of finite differences.
 However, they have the additional property of being conservative: Even if the computed fluxes are rather inaccurate, the sum of fluxes over the spatial domain will vary only according to boundary conditions.
 
-The desired properties of finite volumes hold for any choice of $F_{i+1/2}^{n}$, as long as the resulting method is consistent and stable.
+The desired properties of finite volumes hold for any choice of $F_{i+1/2}^{n}$, as long as the result is consistent and stable.
 Similarly, they hold for any choice of hyperbolic PDE, which is abstracted away as a state space $Q_i$, a pointwise flux $f(Q_i)$ and max wavespeed $\lambda_{max}(Q_i)$.
@@
 Since the system evolves by exchanging fluxes between neighboring cells, it is intuitively necessary that the timestep size be small enough that the fluxes don't travel further than one cell per timestep.
-More precisely, the \emph{numerical domain of dependence} (the set of cells whose old values factor in to the new value of each cell) must contain the \emph{physical domain of dependence}, the region of space from which waves propagated.
-This is captured as a relation between cell size, timestep size, and wave speed, called the Courant number:
+More precisely, the \emph{numerical domain of dependence} (the cells whose old values affect each cell's new value) must contain the \emph{physical domain of dependence}, the spatial region from which waves propagated to a given point, over that same time interval.
+This is expressed via the dimensionless Courant number, which relates cell size, timestep size, and wave speed.
+For a three-point stencil, the Courant number is constrained to:
 
-\[ \nu := \Bigg\lvert\frac{\bar{u}\Delta t}{\Delta x} \Bigg\rvert = \frac{\Delta t}{\Delta x} \lvert \lambda_{max} \rvert  \leq 1 \]
+\begin{equation}
+\label{eq:cfl}
+\nu := \Bigg\lvert\frac{\bar{u}\Delta t}{\Delta x} \Bigg\rvert = \frac{\Delta t}{\Delta x} \lvert \lambda_{max} \rvert  \leq 1
+\end{equation}
 
@@
 The maximum wavespeed calculated from the flux Jacobian is exact only in the small-perturbation case, so an additional correction factor, $\mu \in [0,1]$, scales the timestep down.
-Thus the experimenter has two knobs for balancing speed, accuracy, and stability: The spatial discretization $\Delta x$ and the timestep correction $\mu$.
-
-The CFL condition is encoded on lines 95--98.
+Thus the experimenter has two knobs for balancing speed, accuracy, and stability: the spatial discretization $\Delta x$ and the timestep correction $\mu$.
 
@@
 Different finite-volume methods may be derived from a common skeleton by providing a choice of $\mathcal F$.
-The simplest choice is simply the average of the fluxes at the center of either neighboring cell.
+The simplest choice is merely the average of the fluxes at the center of either neighboring cell.
 
@@
 
-\[ F_{i+1/2} = \frac{1}{2} \Big[ f(Q_{i+1}) + f(Q_{i}) - \frac{\Delta x}{\Delta t}(Q_{i+1} - Q_{i})\Big] \]
+\begin{equation}
+\label{eq:lxf}
+F_{i+1/2} = \frac{1}{2} \Big[ f(Q_{i+1}) + f(Q_{i}) - \frac{\Delta x}{\Delta t}(Q_{i+1} - Q_{i})\Big]
+\end{equation}
+
+The parameter $a := \frac{\Delta x}{\Delta t}$ has multiple interpretations.
+Firstly, it was derived as the $\mu$-scaled max wavespeed over the spatial domain.
+Secondly, it acts as a penalty term: the larger the discontinuity in $Q(x)$, the lower the cellwise flux $F(x)$.
+The following sections will explore its unexpected third interpretation, as a viscosity coefficient.
 
@@
 
-The Lax-Friedrichs flux function is implemented on lines 62--74.
-It was run against two model problems.
-The first used a Gaussian perturbation for its initial condition with reflecting boundary conditions.
-The second used a breaking dam initial condition with outflow boundary conditions.
-Two experiments were run in order to examine the behavior of the knobs $n_{cells}$  and $\mu$ independently.
-These are depicted in Figures~\ref{fig:lxf_gaussian_ncells},~\ref{fig:lxf_breakingdam_ncells}, and~\ref{fig:lxf_breakingdam_mu}.
+The Lax-Friedrichs method was run against two model problems: a Gaussian perturbation with reflecting boundaries, and a breaking dam with outflow boundaries.
+The independent behavior of the knobs $n_{cells}$ and $\mu$ was tested in three experiments.
+The results are depicted in Figures~\ref{fig:lxf_gaussian_ncells},~\ref{fig:lxf_breakingdam_ncells}, and~\ref{fig:lxf_breakingdam_mu}.
 
@@
 
-\[F_{i+1/2} = \frac{1}{2} \Big[f(Q_{i+1} + f(Q_{i})\Big]\]
+\[F_{i+1/2} = \frac{1}{2} \Big[f(Q_{i+1}) + f(Q_{i})\Big]\]
 
@@
 Local-Lax-Friedrichs is a modification which retains the stability benefits of artificial viscosity while decreasing the resulting dissipation.
-In the basic Lax-Friedrichs, the viscosity parameter $a$ is a constant derived from the timestep $\Delta t$, which in turn is derived from the CFL condition.
-Although the timestep is contrained to be constant over the entire spatial domain (local time stepping is not considered in the scope of this paper), the viscosity is not.
-Varying the viscosity over the spatial domain is advantageous: in regions with shocks or other discontinuities, the presence of viscosity improves numerical stability, and in smooth regions, the lack of viscosity leads to a less dissipative and hence more accurate solution.
+In the basic Lax-Friedrichs, the viscosity parameter $a$ is derived from the CFL condition via $\Delta t$.
+Although the timestep must be constant in space (local time stepping is not considered), the viscosity may vary.
+This yields advantages: in regions with shocks or other discontinuities, the presence of viscosity improves numerical stability, and in smooth regions, the lack of viscosity leads to a less dissipative and hence more accurate solution.
 
@@
 
-\[ F_{i-1/2} = \frac{1}{2}[f(Q_{i-1}) + f(Q_i) - a_{i-1/2}(Q_i - Q_{i-1})] \]
+\begin{equation}
+\label{eq:llxf}
+F_{i-1/2} = \frac{1}{2}[f(Q_{i-1}) + f(Q_i) - a_{i-1/2}(Q_i - Q_{i-1})]
+\end{equation}
 
@@
 The Local Lax-Friedrichs method does not entirely break the coupling between dissipation and $\mu$, but it strongly reduces it, as depicted in Figure~\ref{fig:llxf_breakingdam_mu}.
-This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, substituting only the LxF kernel (as implemented in lines 62--74) with the LLxF kernel (as implemented in lines 75--88).
+This experiment used the exact same parameters as Figure~\ref{fig:lxf_breakingdam_mu}, substituting only the LxF kernel (Eq.~\ref{eq:lxf} with the LLxF kernel (Eq.~\ref{eq:llxf}).
 
-\section{LxF vs LLxF Comparison}
+\section{Comparison of LxF vs LLxF}
 
@@
 
-This final section describes the design decisions behind the sample code provided in the appendix.
-The code is broken into chunks corresponding to the theoretical development in the preceding sections.
-It is written in Julia, a language designed for numerical programming.
-The choice of Julia enabled a particularly terse code representation which still maps closely to the theory.
+This final section describes the design decisions behind the implementation.
+It is written in Julia, which permits the code to be both notably terse and closely mapped to the theory.
+The code organization parallels the theoretical development in the preceding sections.
 
 The first design decision is the choice of data representation.
-Julia eschews standard object-oriented idioms in favor of a more functional and data-driven approach.
-The basic composite type is a simple structure, immutable by default, with restricted inheritance and encapsulation.
-
-The state space $Q$ is a mutable structure.
-The spatial domain $q(x) \forall x$ is defined as a one-dimensional array with inner type $Q$.
+Julia eschews standard object-oriented idioms such as inheritance and encapsulation in favor of a more functional and data-driven approach.
+Thus the state space $Q$ is a simple mutable structure.
+The spatial domain $q(x) \forall x$ is defined as a one-dimensional array $QS$ with inner type $Q$.
 The spatial domain has a ghost cell on either side to enforce boundary conditions, so the physical cell indices range over $[2, n_{cells}+1]$.
 
-These additional constraints can be expressed within the type system.
-Firstly, by using \verb!OffsetArray!s instead of regular \verb!Array!s, the indices can be cheaply translated to $[1, n_{cells}]$.
-Secondly, one may parameterize the spatial domain by the number of cells, \verb|QS{ncells}|, e.g.
+These additional constraints could be expressed within the type system.
+Firstly, by using \verb!OffsetArray!s, the indices can be cheaply translated to $[1, n_{cells}]$.
+Secondly, one may parameterize by the number of cells, \verb|QS{ncells}|, e.g.
 by using \verb|StaticArray|s, ensuring that the array size is known at JIT-compile time.
 This allows for array-size correctness checks and compiler optimizations.
-Finally, a custom constructor could encapsulate all of this and enforce the existence of the ghost cells.
-These features may prove helpful on larger projects, though they introduce additional complexity.
+Finally, a custom constructor could encapsulate this and enforce the existence of the ghost cells.
+These features may prove helpful on larger projects, but were omitted here because they introduce more complexity than they resolve.
 
-The choice of an array-of-structures representation is most consistent with the theory, but disregards the useful fact that $Q$ forms a vector space.
-This problem is readily resolved thanks to \emph{multiple dispatch}.
-In Julia, a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.\cite{doi:10.1137/141000671} This means that operations such addition can be extended to custom types in a unified way.
+The choice of an array-of-structures representation is consistent with the theory, but disregards the fact that $Q$ forms a vector space.
+This problem is readily resolved thanks to Julia's \emph{multiple dispatch}, where a function is resolved to a method by examining all of the types in the signature, rather than simply the `owning' object.\cite{doi:10.1137/141000671} This means that operations such addition can be extended to custom types in a unified way.
 Lines 12--26 extend addition and scalar multiplication over the state space type $Q$.
 
-The basic finite volume skeleton is given on lines 38--61.
+ The general finite volume skeleton (Eq.~\ref{eq:fv}) is given on lines 38--61.
 The skeleton accepts abstract strategies for flux computations, timestep computations, initial conditions, and boundary conditions.
 These strategies are implemented as first-class functions rather than as an object heirarchy.
+Strategies for LxF fluxes (Eq.~\ref{eq:lxf}) and LLxF fluxes (Eq.~\ref{eq:llxf}) are implemented on lines 62--74 and 75--88, respectively.
 Lines 91--93 demonstrate a higher-order function which acts like a constructor for timestep strategies.
@@
 
+% The cell-wise fluxes $F_{i+1/2}, F_{i-1/2}$ are represented slightly differently in the code
+
 \section{Conclusion}
@@
 The Lax-Friedrichs method is a simple finite volume technique for solving the shallow water equations.
-Starting from the fundamentals, this paper built up a reference implementation which leverages features of Julia to achieve terseness and clarity not available in languages such as Matlab or C++.
+Starting from the fundamentals, this paper built up a reference implementation which leverages features of Julia to achieve terseness and clarity.
 An abstract finite volume skeleton allows different numerical methods to be swapped in.

### 2018-01-16 paper/paper.tex
Improved captions

@@
  		\includegraphics[scale=0.6]{../plots/lxf_gaussian_ncells.png}%
- 		\caption{LxF performance relative to $n_{cells}$}\label{fig:lxf_gaussian_ncells}%
+ 		\caption{Performance of the Lax-Friedrichs method relative to $n_{cells}$ for a Gaussian perturbation scenario with $t=50s, \mu=0.5$.}\label{fig:lxf_gaussian_ncells}%
  	\end{center}%
@@
  		\includegraphics[scale=0.6]{../plots/lxf_breakingdam_ncells.png}%
- 		\caption{LxF performance relative to $n_{cells}$}\label{fig:lxf_breakingdam_ncells}%
+ 		\caption{Performance of the Lax-Friedrichs method relative to $n_{cells}$ for a breaking dam scenario with $t=50s, \mu=0.5$.}\label{fig:lxf_breakingdam_ncells}%
  	\end{center}%
@@
  		\includegraphics[scale=0.6]{../plots/lxf_breakingdam_mu.png}%
- 		\caption{LxF performance relative to $\mu$}\label{fig:lxf_breakingdam_mu}%
+    \caption{Performance of the Lax-Friedrichs method relative to $\mu$ for a breaking dam scenario with $t = 50s, n_{cells} = 1000$.}\label{fig:lxf_breakingdam_mu}%
  	\end{center}%
@@
         \includegraphics[scale=0.6]{../plots/artificial_viscosity.png}%
-        \caption{Wavespeed over the spatial domain}\label{fig:artificial_viscosity}%
+        \caption{Comparison of height, momentum, and wavespeed over the spatial domain for a Gaussian perturbation with $t = 60s, \mu=1.0, n_{cells} = 100$.}\label{fig:artificial_viscosity}%
     \end{center}%
@@
         \includegraphics[scale=0.6]{../plots/llxf_breakingdam_mu.png}%
-        \caption{Local Lax-Friedrichs, breaking dam scenario. }\label{fig:llxf_breakingdam_mu}%
+        \caption{Performance of Local Lax-Friedrichs relative to $\mu$ for a breaking dam scenario with $t = 50s, n_{cells} = 1000.$ Compare to Figure~\ref{fig:lxf_breakingdam_mu}.}\label{fig:llxf_breakingdam_mu}%
     \end{center}%
@@
         \includegraphics[scale=0.6]{../plots/lxf_llxf_breakingdam.png}%
-        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, breakin dam scenario}\label{fig:cmp_breakingdam}%
+        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs for a breaking dam scenario with $t = 50s, n_{cells} = 200$.}\label{fig:cmp_breakingdam}%
     \end{center}%
@@
         \includegraphics[scale=0.6]{../plots/lxf_llxf_gaussian.png}%
-        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs, gaussian perturbation scenario.}\label{fig:cmp_gaussian}%
+        \caption{Comparison of Lax-Friedrichs vs Local Lax-Friedrichs for a Gaussian perturbation scenario with $t = 200s, n_{cells} = 200$.}\label{fig:cmp_gaussian}%
     \end{center}%
@@
 These strategies are implemented as first-class functions rather than as an object heirarchy.
-Strategies for LxF fluxes (Eq.~\ref{eq:lxf}) and LLxF fluxes (Eq.~\ref{eq:llxf}) are implemented on lines 62--74 and 75--88, respectively.
 Lines 91--93 demonstrate a higher-order function which acts like a constructor for timestep strategies.
